### 2020-10-04 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
v1 First draft of paper, sections Data and Methods and Findings about 70% complete

---
title: An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
author:
  - name: Shaila Jamal
    email: jamals16@mcmaster.ca
    affiliation: McMaster University
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
 The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world. Recent research provides illuminates the nature and magnitude of the changes in mobility, but the evidence is still scant in developing countries. The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh. The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.

journal: "Transport Findings"
date: "`r Sys.Date()`"
bibliography: References.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
always_allow_html: true
---

```{r load-packages, include=FALSE}
#Load the packages needed to read and work with the data
library(circlize)
library(grid)
library(gridExtra)
library(kableExtra)
library(png)
library(readr)
library(tidyverse)
```

```{r read-data, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Read data file, add unique respondent identifier:
mcb <- read_csv("Combined-data-final-with-serial-number.csv") %>% 
  rename(id = `Serial no.`,
         yob = `24.	Please enter your year of birth (e.g. 1990, 1999)`)
```

```{r convert-to-factors-and-clean, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Extract relevant variables and convert to factors as appropriate. A small number of records (4) are deleted because they contain nonsensical year of birth responses 
freq_by_mode <- mcb %>%
  mutate(id = c(1:nrow(mcb)),
         age = 2020 - yob) %>%
  filter(age > 0 & age < 100) %>%
  transmute(id,
            freq_car_before = factor(.[[20]]),
            freq_bus_before = factor(.[[21]]),
            freq_rickshaw_before = factor(.[[22]]),
            freq_cng_before = factor(.[[23]]),
            freq_walk_before = factor(.[[24]]),
            freq_motorcycle_before = factor(.[[25]]),
            freq_bicycle_before = factor(.[[26]]),
            freq_rideshare_before = factor(.[[27]]),
            freq_other_before = factor(.[[28]]),
            freq_car_during = factor(.[[29]]),
            freq_bus_during = factor(.[[30]]),
            freq_rickshaw_during = factor(.[[31]]),
            freq_cng_during = factor(.[[32]]),
            freq_walk_during = factor(.[[33]]),
            freq_motorcycle_during = factor(.[[34]]),
            freq_bicycle_during = factor(.[[35]]),
            freq_rideshare_during = factor(.[[36]]),
            freq_other_during = factor(.[[37]]))
```

```{r frequency-table-pivot-long, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Pivot table to long: each row is a respondent-mode-period response
freq_by_mode_long <- freq_by_mode %>%
  pivot_longer(cols = starts_with("freq_"),
               names_to = "mode_period",
               names_prefix = "freq_",
               values_to = "frequency") %>%
  separate(mode_period, c("Mode", "Period")) %>%
  transmute(id,
            Period = factor(Period,
                            levels = c("before", "during")), 
            Mode = factor(Mode,
                          levels = c("car",
                                     "motorcycle",
                                     "rideshare",
                                     "rickshaw",
                                     "cng",
                                     "bus",
                                     "walk",
                                     "bicycle",
                                     "other")), 
            frequency = factor(frequency, 
                               levels = c("Never",
                                          "Rarely",
                                          "At least once a week",
                                          "2-3 times a week",
                                          "4-5 times a week",
                                          "Almost everyday"),
                               labels = c("Never",
                                          "Rarely",
                                          "Once a week",
                                          "2-3 per week",
                                          "4-5 per week",
                                          "Almost daily"),
                               ordered = TRUE))
```

```{r frequency-table-pivot-wide, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Pivot the table to wide: each row is a respondent
freq_by_mode <- freq_by_mode_long %>%
  pivot_wider(names_from = c("Mode", "Period"),
              names_sep = "_",
              names_prefix = "freq_",
              values_from = "frequency")
```

```{r summarize-frequency, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Summarize cases by period and mode
freq_by_mode_long <- freq_by_mode_long %>% 
  group_by(Period, Mode, frequency) %>%
  summarize(Cases = n())
```

```{r probability-transitions, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Create cross-tabulations and tables of proportions by mode. The cross-tabulation is of the number of cases by period (before COVID-19 and during):

# CAR
car_xtab <- xtabs(~ freq_car_before + freq_car_during, 
                  data = freq_by_mode)

car_ptab <- car_xtab %>%
  prop.table(1)

# MOTORCYCLE
motorcycle_xtab <- xtabs(~ freq_motorcycle_before + freq_motorcycle_during, 
                         data = freq_by_mode)

motorcycle_ptab <- motorcycle_xtab %>%
  prop.table(1)

# RIDESHARE
rideshare_xtab <- xtabs(~ freq_rideshare_before + freq_rideshare_during, 
                        data = freq_by_mode)

rideshare_ptab <- rideshare_xtab %>%
  prop.table(1)

# RICKSHAW
rickshaw_xtab <- xtabs(~ freq_rickshaw_before + freq_rickshaw_during, 
                       data = freq_by_mode)

rickshaw_ptab <- rickshaw_xtab %>%
  prop.table(1)

# CNG
cng_xtab <- xtabs(~ freq_cng_before + freq_cng_during, 
                  data = freq_by_mode)

cng_ptab <- cng_xtab %>%
  prop.table(1)

# BUS
bus_xtab <- xtabs(~ freq_bus_before + freq_bus_during, 
                  data = freq_by_mode)

bus_ptab <- bus_xtab %>%
  prop.table(1)

# WALK
walk_xtab <- xtabs(~ freq_walk_before + freq_walk_during, 
                   data = freq_by_mode)

walk_ptab <- walk_xtab %>%
  prop.table(1)

# BICYCLE
bicycle_xtab <- xtabs(~ freq_bicycle_before + freq_bicycle_during, 
                      data = freq_by_mode)

bicycle_ptab <- bicycle_xtab %>%
  prop.table(1)

# OTHER
other_xtab <- xtabs(~ freq_other_before + freq_other_during, 
                    data = freq_by_mode)

other_ptab <- other_xtab %>%
  prop.table(1)
```

```{r save-data, include=FALSE, eval=FALSE}
# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION

# Save data objects
save(freq_by_mode_long, 
     car_xtab, 
     car_ptab, 
     motorcycle_xtab, 
     motorcycle_ptab, 
     rideshare_xtab, 
     rideshare_ptab, 
     rickshaw_xtab, 
     rickshaw_ptab, 
     cng_xtab, 
     cng_ptab, 
     bus_xtab, 
     bus_ptab, 
     walk_xtab, 
     walk_ptab, 
     bicycle_xtab, 
     bicycle_ptab, 
     other_xtab, 
     other_ptab,
     file = "Trip-Frequency-Bangladesh.RData")
```

```{r load-data-file, include=FALSE}
load("Trip-Frequency-Bangladesh.RData")
```

Research Questions and Hypotheses
==========================

The spread of the COVID-19 pandemic has led to limitations to movement in many countries and regions, either because of lock-down policies or self-censoring by segments of the public. The magnitude of changes in mobility has been studied by recent research, including @DeWeese2020tale and @Molloy2020tracing. While the evidence available indicates that overall there was a reduction in mobility in much  of the world, the changes were uneven depending on the mode of transportation or the purpose of the trip [see @Lock2020cycling; @Paez2020using]. Alas, with few exceptions evidence remains more spotty for developing countries, most of which have large populations segments that are less able to absorb losses in mobility [e.g., @Astroza2020mobility; @Huynh2020culture; @Saha2020lockdown].

The objective of this paper is to investigate changes in the trip-making frequency by different modes of transportation during the COVID-19 emergency in Bangladesh. Using data from a recent survey that asked respondents to report mobility levels before and during the pandemic, we pose the following questions:

- Was there a reduction of mobility in Bangladesh during COVID-19?
- And if so, what forms of transportation were more affected?

This paper is a reproducible research document [see @Brunsdon2020opening]; the code and data necessary to replicate the tables and figures are available in a public repository\footnote{\url{place.holder}}

Methods and Data
============

Data used for this paper come from a survey conducted during **DATES** in **regions/cities** in Bangladesh. **BRIEFLY ADD DETAILS ABOUT SURVEY**.

The survey asked respondents to self-report their trip-making frequency by eight modes of transportation, namely **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **cng auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here). Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_. There are $n=800$ responses in the data set.

To describe changes in the frequency of travel by mode in the transition to the pandemic, we use well-established exploratory data analysis (EDA) techniques.

Findings
===================

Figure \ref{fig:column-plot-cases} shows the number of responses (out of 800) in each trip-making frequency class by mode of transportation. The white bars and gray bars are for travel before and during the pandemic, respectively. Considering travel before the pandemic, travel by rickshaw, bus were relatively common for many respondents (few respondents reported _never_ using these modes). The mode most commonly used on a quotidian basis was walk. In contrast, respondents reported less frequent travel by car, rideshare services, cng auto-rickshaw, motorcycle, and bicycle. During the pandemic we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor. The frequency of trip-making by other modes changed more noticeably: the frequency of travel by rideshare services, rickshaw, cng auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before. The frequency of walking trips also decreased (fewer respondents report walking almost daily), but the reductions in mobility were not so heavily concentrated at the bottom of the scale.

```{r column-plot-cases-before-after, fig.height=6, echo=FALSE, fig.cap="\\label{fig:column-plot-cases}Number of responses by trip-making frequency class and mode, before and during COVID-19"}
freq_by_mode_long %>%
  filter(Mode != "other") %>%
  ggplot(aes(x = frequency, y = Cases, fill = Period)) +
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("before" = "white", "during" = "gray")) +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  facet_wrap(. ~ Mode, nrow = 4)
```

Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of cases in each trip-making frequency class before and during the pandemic. If no changes had occurred, all values would be concentrated on the main diagonal of the matrices. Values in the lower triangular matrix represent a _loss_ of mobility (lower travel frequency), whereas values in the upper triangular matrix are _gains_ (higher travel frequency). The further away a value is from the main diagonal, the greater the loss or gain.

Despite across-the-board losses of mobility, there appears to have been some adaptation that varied by mode of transportation. To illustrate, `r bus_xtab[6, 1]` respondents, or `r paste(round(bus_xtab[6, 1] / sum(bus_xtab[6,]) * 100, 2), "%", sep="")` of those who traveled by bus almost daily before, reported never using it during the pandemic. In contrast, only `r sum(bus_xtab[1, 2:6])` respondents, or `r paste(round(sum(bus_xtab[1, 2:6]) / sum(bus_xtab[,]) * 100, 2), "%", sep="")` of those who never used buses before started doing so during the pandemic. By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so - but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.

```{r table-transitions, echo=FALSE}
rbind(car_xtab, 
      rideshare_xtab,
      rickshaw_xtab, 
      cng_xtab,
      bus_xtab, 
      motorcycle_xtab,
      walk_xtab,
      bicycle_xtab) %>%
  kable("latex",
        booktabs = TRUE,
        caption = "\\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19") %>%
  kable_styling(latex_options = c("scale_down")) %>%
  pack_rows("car", 1, 6) %>%
  pack_rows("rideshare", 7, 12) %>%
  pack_rows("rickshaw", 13, 18) %>%
  pack_rows("cng auto-rickshaw", 19, 24) %>%
  pack_rows("bus", 25, 30) %>%
  pack_rows("motorcycle", 31, 36) %>%
  pack_rows("walk", 37, 42) %>%
  pack_rows("bicycle", 43, 48)
```

To more easily understand the transitions towards different trip-making frequencies, including possible adaptations by mode, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.

Figures \ref{fig:circular-plot-1} to \ref{fig:circular-plot-4} present these plots. Each of the trip-making frequency sectors on the left hemisphere of the circle represent 100% of responses _before_ the pandemic. The size of the links is proportional to the probability $p_{ij}$ of transitioning from frequency class $i$ before to the right hemisphere are proportional to the transition probabilities to each frequency _during_ the pandemic. There are three transparency levels for the links: solid colors are for $p_{ij}>2/3$, intermediate transparency is for $1/3 < p_{ij} \le 2/3$, and the more transparent links are for $p_{ij}\le 1/3$.

From Figure \ref{fig:circular-plot-1} we see that the probability of traveling less frequently for those who frequently traveled by car before is high, but their probability of not using this mode at all during the pandemic is quite small. The probability of traveling more frequently by car for those who originally never or rarely used this mode remained is low. In contrast, we see that the probabilities of never ridesharing during the pandemic are high irrespective of the initial level of use of this mode of transportation.

The probabilities of change in trip frequency by rickshaw and and cng auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}), although the probabilities of being less mobile by cng auto-rickshaw are greater: between 1/3 and 2/3 of respondents who used this mode almost daily, stopped using it during the pandemic. Very rarely there was an increase in mobility by these modes.

After ridesharers, bus users (Figure \ref{fig:circular-plot-3}) were the most likely to stop using it, irrespective of their initial level of use of this mode, while the probabilities of using it _more_ frequently are extremely small.

```{r reorder-probability-transition-matrices, include=FALSE}
# For visualization purposes using `circlize`, I want to reverse the order of the columns of the table of proportions. In this way, "Never" and "Never" are adjacent, and "Almost daily" and "Almost daily" are also adjacent in the circular plot:

car_ptab2 <- car_ptab[, c(6, 5, 4, 3, 2, 1)]
motorcycle_ptab2 <- motorcycle_ptab[, c(6, 5, 4, 3, 2, 1)]
rideshare_ptab2 <- rideshare_ptab[, c(6, 5, 4, 3, 2, 1)]
rickshaw_ptab2 <- rickshaw_ptab[, c(6, 5, 4, 3, 2, 1)]
cng_ptab2 <- cng_ptab[, c(6, 5, 4, 3, 2, 1)]
bus_ptab2 <- bus_ptab[, c(6, 5, 4, 3, 2, 1)]
walk_ptab2 <- walk_ptab[, c(6, 5, 4, 3, 2, 1)]
bicycle_ptab2 <- bicycle_ptab[, c(6, 5, 4, 3, 2, 1)]
other_ptab2 <- other_ptab[, c(6, 5, 4, 3, 2, 1)]
```

```{r rename-states-in-probability-transition-matrices, include=FALSE}
# Copy the states and rename the rows and columns of the table so that `circlize` knows that the states are _not_ the same, but rather correspond to _before_ and _during_:

all_states <- rownames(bicycle_ptab2) # Names of the states, that is, the frequency categories
n_states <- length(all_states) # Numer of states

# CAR
rownames(car_ptab2) = paste0("B_", all_states)
colnames(car_ptab2) = paste0("D_", rev(all_states))

# MOTORCYCLE
rownames(motorcycle_ptab2) = paste0("B_", all_states)
colnames(motorcycle_ptab2) = paste0("D_", rev(all_states))

# RIDESHARE
rownames(rideshare_ptab2) = paste0("B_", all_states)
colnames(rideshare_ptab2) = paste0("D_", rev(all_states))

# RICKSHAW
rownames(rickshaw_ptab2) = paste0("B_", all_states)
colnames(rickshaw_ptab2) = paste0("D_", rev(all_states))

# CNG
rownames(cng_ptab2) = paste0("B_", all_states)
colnames(cng_ptab2) = paste0("D_", rev(all_states))

# BUS
rownames(bus_ptab2) = paste0("B_", all_states)
colnames(bus_ptab2) = paste0("D_", rev(all_states))

# WALK
rownames(walk_ptab2) = paste0("B_", all_states)
colnames(walk_ptab2) = paste0("D_", rev(all_states))

# BICYCLE
rownames(bicycle_ptab2) <- paste0("B_", all_states)
colnames(bicycle_ptab2) <- paste0("D_", rev(all_states))

# OTHER
rownames(other_ptab2) = paste0("B_", all_states)
colnames(other_ptab2) = paste0("D_", rev(all_states))
```

```{r color-matrix-for-links, include=FALSE}
#Create a colors matrix for the links in the circular plot:
state_col <- c("Never" = "#2C3E50",
               "Rarely" = "#1560BD",
               "Once a week" = "#3498DB",
               "2-3 per week" = "#F1C40F",
               "4-5 per week" = "#E67E22",
               "Almost daily" = "#C0392B")

# one for rows and one for columns
state_col2 <- c(state_col, rev(state_col))
names(state_col2) <- c(rownames(bicycle_ptab2), colnames(bicycle_ptab2))

colmat <- rep(state_col2[rownames(bicycle_ptab2)], n_states)
colmat <- rgb(t(col2rgb(colmat)), maxColorValue = 255)
```

```{r grouping-variable-for-period, include=FALSE}
#Create a grouping variable for the chord diagram, so that sectors are grouped by their state as before and during:
group = structure(c("Before", "Before", "Before", "Before", "Before", "Before", "During", "During", "During", "During", "During", "During"), names = names(state_col2))
```

```{r function-for-circular-plots, include=FALSE}
# Define a function to create circular plots. The inputs are a matrix of transition probabilities, the name of the mode for the title, and the number of lines to adjust the position of the title
circle.plot <- function(ptab2 = ptab2, mode.title = "mode", nlines = 0){
  factors <- factor(c("Before COVID-19", "During COVID-19"))
  circos.clear()
  circos.par(start.degree = 270, 
             points.overflow.warning = FALSE,
             track.height = 0.2)
  circos.initialize(factors = factors, xlim = c(0, 1))
  circos.track(ylim = c(0, 1), 
               panel.fun = function(x, y) {
                 circos.text(CELL_META$xcenter, 
                             CELL_META$ycenter + uy(1.8, "mm"),
                             CELL_META$sector.index,
                             cex = 0.8,
                             facing = "bending.inside",
                             niceFacing = FALSE)
               })

  title(paste0("Frequency of travel by ", mode.title, sep = ""),
        line = nlines,
        cex.main = 0.8)

  par(new = TRUE) # <- magic

  # Clear parameters for circular plot
  circos.clear()

  # Set the start degree so that the circular plot is symmetric on the vertical axis
  circos.par("canvas.xlim" = c(-1, 1), 
             "canvas.ylim" = c(-1, 1), 
             start.degree = 268, 
             points.overflow.warning = FALSE)

  # Transition probabilities lower than 2/3 and 1/3 are more transparent (this is done by pasting a transparency value to the color)
  # Hexadecimal transparency values can be found here: https://gist.github.com/lopspower/03fb1cc0ac9f32ef38f4
  # Make a copy of the color matrix to modify. The changes are specific to each mode, so we do not want to change the original matrix of colors
  colmat_m <- colmat
  colmat_m[ptab2 > 2/3] <- paste0(colmat_m[ptab2 > 2/3], "FF")
  colmat_m[ptab2 <= 2/3 & ptab2 > 1/3] <- paste0(colmat_m[ptab2 <= 2/3 & ptab2 > 1/3], "A8")
  colmat_m[ptab2 <= 1/3] <- paste0(colmat_m[ptab2 <= 1/3], "33")
  dim(colmat_m) = dim(ptab2)

  chordDiagram(ptab2, 
               preAllocateTracks = list(list(track.height = 0.1),
                                        list(track.height = 0.05)),
               col = colmat_m, 
               grid.col = state_col2,
               directional = TRUE,
               diffHeight = mm_h(2),
               #link.arr.type = "big.arrow",
               group = group,
               annotationTrack = "grid",
               big.gap = 5)

  circos.track(track.index = 2,
               ylim = c(0, 1),
               panel.fun = function(x, y){
                 if(abs(CELL_META$cell.start.degree - CELL_META$cell.end.degree) > 3){
                   sn = CELL_META$sector.index
                   i_state = (gsub("(B|D)_", "", sn))
                   circos.text(CELL_META$xcenter, 
                               CELL_META$ycenter + uy(0, "mm"), 
                               i_state, 
                               col = "black", 
                               font = 1, 
                               cex = 0.6, 
                               adj = c(0.5, 0.5), 
                               facing = "bending.inside",
                               niceFacing = TRUE)}
               },
               bg.border = NA)
}
```

```{r example-of-circular-plot, echo=FALSE, eval=FALSE, fig.height=8}
# NOT RUN - THIS IS AN EXAMPLE OF THE CODE USED TO CREATE A SINGLE CIRCULAR PLOT
circle.plot(bus_ptab2, "bus")
```

```{r circular-plots-transition-probabilities-1, echo = FALSE, fig.cap="\\label{fig:circular-plot-1}Transition probabilities in trip frequency from before to during COVID-19: car and rideshare"}
# Rows and columns to arrange several circular plots
par(mfrow = c(1, 2),
    mar = c(0, 0, 0, 0),
    oma = c(0, 0, 0, 0))

# CAR
circle.plot(car_ptab2, "car", nlines = -3)

# RIDESHARE
circle.plot(rideshare_ptab2, "rideshare", nlines = -3)
```

```{r circular-plots-transition-probabilities-2, echo = FALSE, fig.cap="\\label{fig:circular-plot-2}Transition probabilities in trip frequency from before to during COVID-19: rickshaw and cng auto-rickshaw"}
# Rows and columns to arrange several circular plots
par(mfrow = c(1, 2),
    mar = c(0, 0, 0, 0),
    oma = c(0, 0, 0, 0))

# RICKSHAW
circle.plot(rickshaw_ptab2, "rickshaw", nlines = -3)

# CNG
circle.plot(cng_ptab2, "cng auto-rickshaw", nlines = -3)
```

```{r circular-plots-transition-probabilities-3, echo = FALSE, fig.cap="\\label{fig:circular-plot-3}Transition probabilities in trip frequency from before to during COVID-19: motorcycle and bus"}
# Rows and columns to arrange several circular plots
par(mfrow = c(1, 2),
    mar = c(0, 0, 0, 0),
    oma = c(0, 0, 0, 0))

# BUS
circle.plot(bus_ptab2, "bus", nlines = -3)

# MOTORCYCLE
circle.plot(motorcycle_ptab2, "motorcycle", nlines = -3)
```

```{r circular-plots-transition-probabilities-4, echo = FALSE, fig.cap="\\label{fig:circular-plot-4}Transition probabilities in trip frequency from before to during COVID-19: bicycle and walk"}
# Rows and columns to arrange several circular plots
par(mfrow = c(1, 2),
    mar = c(0, 0, 0, 0),
    oma = c(0, 0, 0, 0))

# WALK
circle.plot(walk_ptab2, "walk", nlines = -3)

# BICYCLE
circle.plot(bicycle_ptab2, "bicycle", nlines = -3)
```

References {#references .unnumbered}
==========

### 2020-10-04 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.tex
v1 First draft of paper, sections Data and Methods and Findings about 70% complete

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Transport Findings} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}
\begin{frontmatter}

  \title{An exploratory analysis of changes in trip-making frequency by mode
during the COVID-19 emergency in Bangladesh}
    \author[McMaster University]{Shaila Jamal}
   \ead{jamals16@mcmaster.ca} 
    \author[McMaster University]{Antonio Paez\corref{1}}
   \ead{paezha@mcmaster.ca} 
      \address[McMaster University]{School of Earth, Environment and Society, McMaster University, Hamilton,
ON, L8S 4K1, Canada}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The COVID-19 pandemic has had a profound impact on mobility in every
  country and region around the world. Recent research provides
  illuminates the nature and magnitude of the changes in mobility, but the
  evidence is still scant in developing countries. The objective of this
  paper is to present an exploratory analysis of the changes in the
  frequency of trip-making by mode during the COVID-19 emergency in
  Bangladesh. The results confirm an overall loss of mobility in the form
  of reduced trip-making frequency by all modes, but the changes are not
  uniform across modes.
  \end{abstract}

 \end{frontmatter}

\hypertarget{research-questions-and-hypotheses}{%
\section{Research Questions and
Hypotheses}\label{research-questions-and-hypotheses}}

The spread of the COVID-19 pandemic has led to limitations to movement
in many countries and regions, either because of lock-down policies or
self-censoring by segments of the public. The magnitude of changes in
mobility has been studied by recent research, including DeWeese et al.
(2020) and Molloy et al. (2020). While the evidence available indicates
that overall there was a reduction in mobility in much of the world, the
changes were uneven depending on the mode of transportation or the
purpose of the trip (see Lock, 2020; Paez, 2020). Alas, with few
exceptions evidence remains more spotty for developing countries, most
of which have large populations segments that are less able to absorb
losses in mobility (e.g., Astroza et al., 2020; Huynh, 2020; Saha et
al., 2020).

The objective of this paper is to investigate changes in the trip-making
frequency by different modes of transportation during the COVID-19
emergency in Bangladesh. Using data from a recent survey that asked
respondents to report mobility levels before and during the pandemic, we
pose the following questions:

\begin{itemize}
\tightlist
\item
  Was there a reduction of mobility in Bangladesh during COVID-19?
\item
  And if so, what forms of transportation were more affected?
\end{itemize}

This paper is a reproducible research document (see Brunsdon and Comber,
2020); the code and data necessary to replicate the tables and figures
are available in a public repository\footnote{\url{place.holder}}

\hypertarget{methods-and-data}{%
\section{Methods and Data}\label{methods-and-data}}

Data used for this paper come from a survey conducted during
\textbf{DATES} in \textbf{regions/cities} in Bangladesh. \textbf{BRIEFLY
ADD DETAILS ABOUT SURVEY}.

The survey asked respondents to self-report their trip-making frequency
by eight modes of transportation, namely \textbf{car},
\textbf{ridesharing} (e.g., Uber, Pathao), \textbf{rickshaw},
\textbf{cng auto-rickshaw} (a rickshaw-like vehicle powered by
compressed natural gas), \textbf{bus}, motorcycle/scooter (hereafter
just \textbf{motorcycle}), \textbf{walking}, and \textbf{bicycle} (there
was an additional catch-all category \textbf{other} which we ignore
here). Participants in the survey used the following levels to report
their frequency of traveling by each mode both before and during
COVID-19: \emph{Never}, \emph{Rarely}, \emph{Once a week}, \emph{2-3
trips per week}, \emph{4-5 trips per week}, \emph{Almost daily}. There
are \(n=800\) responses in the data set.

To describe changes in the frequency of travel by mode in the transition
to the pandemic, we use well-established exploratory data analysis (EDA)
techniques.

\hypertarget{findings}{%
\section{Findings}\label{findings}}

Figure \ref{fig:column-plot-cases} shows the number of responses (out of
800) in each trip-making frequency class by mode of transportation. The
white bars and gray bars are for travel before and during the pandemic,
respectively. Considering travel before the pandemic, travel by
rickshaw, bus were relatively common for many respondents (few
respondents reported \emph{never} using these modes). The mode most
commonly used on a quotidian basis was walk. In contrast, respondents
reported less frequent travel by car, rideshare services, cng
auto-rickshaw, motorcycle, and bicycle. During the pandemic we see that
while there were reductions in mobility by car, motorcycle, and bicycle
(with more respondents reporting never traveling by these modes), the
changes were relatively minor. The frequency of trip-making by other
modes changed more noticeably: the frequency of travel by rideshare
services, rickshaw, cng auto-rickshaw, and bus collapsed, with vastly
more respondents reporting never using these modes during the pandemic
than before. The frequency of walking trips also decreased (fewer
respondents report walking almost daily), but the reductions in mobility
were not so heavily concentrated at the bottom of the scale.

\begin{figure}
\centering
\includegraphics{Frequency-of-Travel-by-Mode-COVID-19-Bangladesh_files/figure-latex/column-plot-cases-before-after-1.pdf}
\caption{\label{fig:column-plot-cases}Number of responses by trip-making
frequency class and mode, before and during COVID-19}
\end{figure}

Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of
cases in each trip-making frequency class before and during the
pandemic. If no changes had occurred, all values would be concentrated
on the main diagonal of the matrices. Values in the lower triangular
matrix represent a \emph{loss} of mobility (lower travel frequency),
whereas values in the upper triangular matrix are \emph{gains} (higher
travel frequency). The further away a value is from the main diagonal,
the greater the loss or gain.

Despite across-the-board losses of mobility, there appears to have been
some adaptation that varied by mode of transportation. To illustrate,
103 respondents, or 65.61\% of those who traveled by bus almost daily
before, reported never using it during the pandemic. In contrast, only
12 respondents, or 1.5\% of those who never used buses before started
doing so during the pandemic. By way of comparison, 24.14\% of
respondents who cycled almost daily before the pandemic stopped doing so
- but 4.88\% who never cycled before started doing so during the
pandemic.

\begin{table}

\caption{\label{tab:table-transitions}\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrr}
\toprule
  & Never & Rarely & Once a week & 2-3 per week & 4-5 per week & Almost daily\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{car}}\\
\hspace{1em}Never & 267 & 20 & 5 & 1 & 1 & 2\\
\hspace{1em}Rarely & 142 & 99 & 21 & 12 & 9 & 3\\
\hspace{1em}Once a week & 11 & 23 & 19 & 7 & 2 & 3\\
\hspace{1em}2-3 per week & 10 & 23 & 7 & 6 & 8 & 2\\
\hspace{1em}4-5 per week & 4 & 9 & 5 & 6 & 12 & 4\\
\hspace{1em}Almost daily & 6 & 15 & 6 & 10 & 0 & 20\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{rideshare}}\\
\hspace{1em}Never & 254 & 13 & 6 & 4 & 0 & 0\\
\hspace{1em}Rarely & 192 & 44 & 12 & 1 & 1 & 1\\
\hspace{1em}Once a week & 60 & 20 & 31 & 3 & 1 & 0\\
\hspace{1em}2-3 per week & 49 & 12 & 9 & 7 & 2 & 0\\
\hspace{1em}4-5 per week & 28 & 10 & 3 & 5 & 6 & 2\\
\hspace{1em}Almost daily & 14 & 3 & 4 & 1 & 0 & 2\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{rickshaw}}\\
\hspace{1em}Never & 51 & 9 & 6 & 0 & 1 & 0\\
\hspace{1em}Rarely & 64 & 70 & 18 & 4 & 3 & 1\\
\hspace{1em}Once a week & 36 & 38 & 45 & 9 & 3 & 3\\
\hspace{1em}2-3 per week & 34 & 36 & 18 & 29 & 6 & 4\\
\hspace{1em}4-5 per week & 21 & 30 & 7 & 16 & 15 & 8\\
\hspace{1em}Almost daily & 45 & 65 & 30 & 18 & 16 & 41\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{cng auto-rickshaw}}\\
\hspace{1em}Never & 176 & 14 & 3 & 1 & 0 & 0\\
\hspace{1em}Rarely & 213 & 95 & 14 & 4 & 1 & 3\\
\hspace{1em}Once a week & 36 & 31 & 38 & 9 & 2 & 3\\
\hspace{1em}2-3 per week & 29 & 18 & 16 & 14 & 8 & 3\\
\hspace{1em}4-5 per week & 13 & 6 & 5 & 9 & 14 & 1\\
\hspace{1em}Almost daily & 8 & 7 & 2 & 0 & 0 & 4\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{bus}}\\
\hspace{1em}Never & 116 & 9 & 3 & 0 & 0 & 0\\
\hspace{1em}Rarely & 155 & 46 & 9 & 0 & 0 & 1\\
\hspace{1em}Once a week & 73 & 32 & 22 & 4 & 0 & 0\\
\hspace{1em}2-3 per week & 43 & 19 & 11 & 7 & 5 & 0\\
\hspace{1em}4-5 per week & 60 & 6 & 5 & 4 & 12 & 1\\
\hspace{1em}Almost daily & 103 & 21 & 10 & 11 & 2 & 10\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{motorcycle}}\\
\hspace{1em}Never & 467 & 22 & 12 & 3 & 1 & 8\\
\hspace{1em}Rarely & 44 & 38 & 13 & 10 & 0 & 4\\
\hspace{1em}Once a week & 11 & 11 & 34 & 3 & 3 & 3\\
\hspace{1em}2-3 per week & 13 & 6 & 6 & 15 & 4 & 2\\
\hspace{1em}4-5 per week & 6 & 4 & 0 & 4 & 14 & 3\\
\hspace{1em}Almost daily & 1 & 3 & 2 & 4 & 3 & 23\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{walk}}\\
\hspace{1em}Never & 17 & 7 & 3 & 4 & 0 & 2\\
\hspace{1em}Rarely & 26 & 53 & 13 & 6 & 2 & 4\\
\hspace{1em}Once a week & 12 & 19 & 54 & 9 & 4 & 2\\
\hspace{1em}2-3 per week & 10 & 14 & 12 & 26 & 13 & 4\\
\hspace{1em}4-5 per week & 7 & 3 & 7 & 11 & 16 & 8\\
\hspace{1em}Almost daily & 46 & 73 & 40 & 49 & 27 & 197\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{bicycle}}\\
\hspace{1em}Never & 504 & 22 & 7 & 4 & 2 & 4\\
\hspace{1em}Rarely & 59 & 43 & 20 & 4 & 2 & 4\\
\hspace{1em}Once a week & 14 & 13 & 21 & 3 & 2 & 1\\
\hspace{1em}2-3 per week & 3 & 4 & 5 & 7 & 2 & 0\\
\hspace{1em}4-5 per week & 2 & 2 & 4 & 1 & 9 & 3\\
\hspace{1em}Almost daily & 7 & 1 & 1 & 1 & 3 & 16\\
\bottomrule
\end{tabular}}
\end{table}

To more easily understand the transitions towards different trip-making
frequencies, including possible adaptations by mode, we convert the
cross-tabulations to probability transition matrices, which we then
visualize using circular plots.

Figures \ref{fig:circular-plot-1} to \ref{fig:circular-plot-4} present
these plots. Each of the trip-making frequency sectors on the left
hemisphere of the circle represent 100\% of responses \emph{before} the
pandemic. The size of the links is proportional to the probability
\(p_{ij}\) of transitioning from frequency class \(i\) before to the
right hemisphere are proportional to the transition probabilities to
each frequency \emph{during} the pandemic. There are three transparency
levels for the links: solid colors are for \(p_{ij}>2/3\), intermediate
transparency is for \(1/3 < p_{ij} \le 2/3\), and the more transparent
links are for \(p_{ij}\le 1/3\).

From Figure \ref{fig:circular-plot-1} we see that the probability of
traveling less frequently for those who frequently traveled by car
before is high, but their probability of not using this mode at all
during the pandemic is quite small. The probability of traveling more
frequently by car for those who originally never or rarely used this
mode remained is low. In contrast, we see that the probabilities of
never ridesharing during the pandemic are high irrespective of the
initial level of use of this mode of transportation.

The probabilities of change in trip frequency by rickshaw and and cng
auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}),
although the probabilities of being less mobile by cng auto-rickshaw are
greater: between 1/3 and 2/3 of respondents who used this mode almost
daily, stopped using it during the pandemic. Very rarely there was an
increase in mobility by these modes.

After ridesharers, bus users (Figure \ref{fig:circular-plot-3}) were the
most likely to stop using it, irrespective of their initial level of use
of this mode, while the probabilities of using it \emph{more} frequently
are extremely small.

\begin{figure}
\centering
\includegraphics{Frequency-of-Travel-by-Mode-COVID-19-Bangladesh_files/figure-latex/circular-plots-transition-probabilities-1-1.pdf}
\caption{\label{fig:circular-plot-1}Transition probabilities in trip
frequency from before to during COVID-19: car and rideshare}
\end{figure}

\begin{figure}
\centering
\includegraphics{Frequency-of-Travel-by-Mode-COVID-19-Bangladesh_files/figure-latex/circular-plots-transition-probabilities-2-1.pdf}
\caption{\label{fig:circular-plot-2}Transition probabilities in trip
frequency from before to during COVID-19: rickshaw and cng
auto-rickshaw}
\end{figure}

\begin{figure}
\centering
\includegraphics{Frequency-of-Travel-by-Mode-COVID-19-Bangladesh_files/figure-latex/circular-plots-transition-probabilities-3-1.pdf}
\caption{\label{fig:circular-plot-3}Transition probabilities in trip
frequency from before to during COVID-19: motorcycle and bus}
\end{figure}

\begin{figure}
\centering
\includegraphics{Frequency-of-Travel-by-Mode-COVID-19-Bangladesh_files/figure-latex/circular-plots-transition-probabilities-4-1.pdf}
\caption{\label{fig:circular-plot-4}Transition probabilities in trip
frequency from before to during COVID-19: bicycle and walk}
\end{figure}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Astroza2020mobility}{}%
Astroza, S., Tirachini, A., Hurtubia, R., Carrasco, J.A., Guevara, A.,
Munizaga, M., Figueroa, M., Torres, V., 2020. Mobility changes,
teleworking, and remote communication during the covid-19 pandemic in
chile. Transport Findings.
doi:\href{https://doi.org/10.32866/001c.13489}{10.32866/001c.13489}

\leavevmode\hypertarget{ref-Brunsdon2020opening}{}%
Brunsdon, C., Comber, A., 2020. Opening practice: Supporting
reproducibility and critical spatial data science. Journal of
Geographical Systems 1--20.

\leavevmode\hypertarget{ref-DeWeese2020tale}{}%
DeWeese, J., Hawa, L., Demyk, H., Davey, Z., Belikow, A., El-geneidy,
A., 2020. A tale of 40 cities: A preliminary analysis of equity impacts
of covid-19 service adjustments across north america. Tansport Findings.
doi:\href{https://doi.org/10.32866/001c.13395}{10.32866/001c.13395}

\leavevmode\hypertarget{ref-Huynh2020culture}{}%
Huynh, T.L.D., 2020. Does culture matter social distancing under the
covid-19 pandemic? Safety Science 130, 7.
doi:\href{https://doi.org/10.1016/j.ssci.2020.104872}{10.1016/j.ssci.2020.104872}

\leavevmode\hypertarget{ref-Lock2020cycling}{}%
Lock, O., 2020. Cycling behaviour changes as a result of covid-19: A
survey of users in sydney, australia. Transport Findings.
doi:\href{https://doi.org/10.32866/001c.13405}{10.32866/001c.13405}

\leavevmode\hypertarget{ref-Molloy2020tracing}{}%
Molloy, J., Tchervenkov, C., Hintermann, B., Axhausen, K.W., 2020.
Tracing the sars-cov-2 impact: The first month in switzerland. Transport
Findings.
doi:\href{https://doi.org/10.32866/001c.12903}{10.32866/001c.12903}

\leavevmode\hypertarget{ref-Paez2020using}{}%
Paez, A., 2020. Using google community mobility reports to investigate
the incidence of covid-19 in the united states. Transport Findings.
doi:\href{https://doi.org/10.32866/001c.12976}{10.32866/001c.12976}

\leavevmode\hypertarget{ref-Saha2020lockdown}{}%
Saha, J., Barman, B., Chouhan, P., 2020. Lockdown for covid-19 and its
impact on community mobility in india: An analysis of the covid-19
community mobility reports, 2020. Children and Youth Services Review
116, 14.
doi:\href{https://doi.org/10.1016/j.childyouth.2020.105160}{10.1016/j.childyouth.2020.105160}

\end{document}

### 2020-10-05 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
Survey description added. proofread by shaila

@@
  The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
-Recent research provides illuminates the nature and magnitude of the changes in mobility, but the evidence is still scant in developing countries.
+Recent research illuminates the nature and magnitude of the changes in mobility, but the evidence is still scant in developing countries.
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.
@@
 
-# CNG
+# cng
 cng_xtab <- xtabs(~ freq_cng_before + freq_cng_during, 
@@
 The magnitude of changes in mobility has been studied by recent research, including @DeWeese2020tale and @Molloy2020tracing.
-While the evidence available indicates that overall there was a reduction in mobility in much  of the world, the changes were uneven depending on the mode of transportation or the purpose of the trip [see @Lock2020cycling; @Paez2020using].
-Alas, with few exceptions evidence remains more spotty for developing countries, most of which have large populations segments that are less able to absorb losses in mobility [e.g., @Astroza2020mobility; @Huynh2020culture; @Saha2020lockdown].
+While the evidence available indicates that although the overall mobility has reduced in much  of the world, the changes were uneven depending on the mode of transportation or the purpose of the trip [see @Lock2020cycling; @Paez2020using].
+Unfortunately, with few exceptions, evidence remains more spotty for developing countries, most of which have large population segments that are less able to absorb losses in mobility [e.g., @Astroza2020mobility; @Huynh2020culture; @Saha2020lockdown].
 
@@
 
-Data used for this paper come from a survey conducted during **DATES** in **regions/cities** in Bangladesh. **BRIEFLY ADD DETAILS ABOUT SURVEY**.
+Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lockdown and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
+The survey was available both in Bengali and English and disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
+Including socio-economic characteristics, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lockdown and post-pandemic situation.
 
-The survey asked respondents to self-report their trip-making frequency by eight modes of transportation, namely **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **cng auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
+The survey asked respondents to self-report their trip-making frequency by eight modes of transportation, namely **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
 Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
@@
 
-To describe changes in the frequency of travel by mode in the transition to the pandemic, we use well-established exploratory data analysis (EDA) techniques.
+To describe changes in travel frequency by mode in the transition to the pandemic, we use well-established exploratory data analysis (EDA) techniques.
 
@@
 The white bars and gray bars are for travel before and during the pandemic, respectively.
-Considering travel before the pandemic, travel by rickshaw, bus were relatively common for many respondents (few respondents reported _never_ using these modes).
+Considering travel before the pandemic, travel by rickshaw, and bus were relatively common for many respondents (few respondents reported _never_ using these modes).
 The mode most commonly used on a quotidian basis was walk.
-In contrast, respondents reported less frequent travel by car, rideshare services, cng auto-rickshaw, motorcycle, and bicycle.
-During the pandemic we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor.
-The frequency of trip-making by other modes changed more noticeably: the frequency of travel by rideshare services, rickshaw, cng auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
+In contrast, respondents reported less frequent travel by car, ridesharing services, CNG auto-rickshaw, motorcycle, and bicycle.
+During the pandemic, we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor.
+The frequency of trip-making by other modes changed more noticeably: the frequency of travel by ridesharing services, rickshaw, CNG auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
 The frequency of walking trips also decreased (fewer respondents report walking almost daily), but the reductions in mobility were not so heavily concentrated at the bottom of the scale.
@@
 
-To more easily understand the transitions towards different trip-making frequencies, including possible adaptations by mode, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
+To understand the transitions towards different trip-making frequencies more clearly, including possible adaptations by mode, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
 
 Figures \ref{fig:circular-plot-1} to \ref{fig:circular-plot-4} present these plots.
-Each of the trip-making frequency sectors on the left hemisphere of the circle represent 100% of responses _before_ the pandemic.
-The size of the links is proportional to the probability $p_{ij}$ of transitioning from frequency class $i$ before to the right hemisphere are proportional to the transition probabilities to each frequency _during_ the pandemic.
+Each of the trip-making frequency sectors on the left hemisphere of the circle represents 100% of responses _before_ the pandemic.
+The links' size is proportional to the probability $p_{ij}$ of transitioning from frequency class $i$ before to the right hemisphere are proportional to the transition probabilities to each frequency _during_ the pandemic.
 There are three transparency levels for the links: solid colors are for $p_{ij}>2/3$, intermediate transparency is for $1/3 < p_{ij} \le 2/3$, and the more transparent links are for $p_{ij}\le 1/3$.
 
-From Figure \ref{fig:circular-plot-1} we see that the probability of traveling less frequently for those who frequently traveled by car before is high, but their probability of not using this mode at all during the pandemic is quite small.
+From Figure \ref{fig:circular-plot-1}, we see that the probability of traveling less frequently for those who frequently traveled by car before is high, but their probability of not using this mode at all during the pandemic is quite small.
 The probability of traveling more frequently by car for those who originally never or rarely used this mode remained is low.
@@
 
-The probabilities of change in trip frequency by rickshaw and and cng auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}), although the probabilities of being less mobile by cng auto-rickshaw are greater: between 1/3 and 2/3 of respondents who used this mode almost daily, stopped using it during the pandemic.
+The probabilities of change in trip frequency by rickshaw and CNG auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}), although the probabilities of being less mobile by CNG auto-rickshaw are greater: between 1/3 and 2/3 of respondents who used this mode almost daily, stopped using it during the pandemic.
 Very rarely there was an increase in mobility by these modes.
 
-After ridesharers, bus users (Figure \ref{fig:circular-plot-3}) were the most likely to stop using it, irrespective of their initial level of use of this mode, while the probabilities of using it _more_ frequently are extremely small.
+After rideshare users, bus users (Figure \ref{fig:circular-plot-3}) were the most likely to stop using it, irrespective of their initial level of use of this mode, while the probabilities of using it _more_ frequently are extremely small.
 

### 2020-10-05 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
Survey description added. proofread by shaila

@@
 
-To understand the transitions towards different trip-making frequencies more clearly, including possible adaptations by mode, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
+To more clearly understand the transitions towards different trip-making frequencies, including possible adaptations, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
 

### 2020-10-06 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
v1 Paper submitted to Transport Findings

@@
  The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
-Recent research illuminates the nature and magnitude of the changes in mobility, but the evidence is still scant in developing countries.
+Recent helps to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.
 The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
+In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw.
 
@@
 always_allow_html: true
+
+header-includes:
+ \usepackage{float}
+ \floatplacement{figure}{H}
 ---
@@
 
-Research Questions and Hypotheses
-==========================
-
-The spread of the COVID-19 pandemic has led to limitations to movement in many countries and regions, either because of lock-down policies or self-censoring by segments of the public.
-The magnitude of changes in mobility has been studied by recent research, including @DeWeese2020tale and @Molloy2020tracing.
-While the evidence available indicates that although the overall mobility has reduced in much  of the world, the changes were uneven depending on the mode of transportation or the purpose of the trip [see @Lock2020cycling; @Paez2020using].
-Unfortunately, with few exceptions, evidence remains more spotty for developing countries, most of which have large population segments that are less able to absorb losses in mobility [e.g., @Astroza2020mobility; @Huynh2020culture; @Saha2020lockdown].
-
-The objective of this paper is to investigate changes in the trip-making frequency by different modes of transportation during the COVID-19 emergency in Bangladesh.
-Using data from a recent survey that asked respondents to report mobility levels before and during the pandemic, we pose the following questions:
-
-- Was there a reduction of mobility in Bangladesh during COVID-19?
-- And if so, what forms of transportation were more affected?
-
-This paper is a reproducible research document [see @Brunsdon2020opening]; the code and data necessary to replicate the tables and figures are available in a public repository\footnote{\url{place.holder}}
-
-Methods and Data
-============
-
-Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lockdown and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
-The survey was available both in Bengali and English and disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
-Including socio-economic characteristics, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lockdown and post-pandemic situation.
-
-The survey asked respondents to self-report their trip-making frequency by eight modes of transportation, namely **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
-Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
-There are $n=800$ responses in the data set.
-
-To describe changes in travel frequency by mode in the transition to the pandemic, we use well-established exploratory data analysis (EDA) techniques.
-
-Findings
-===================
-
-Figure \ref{fig:column-plot-cases} shows the number of responses (out of 800) in each trip-making frequency class by mode of transportation.
-The white bars and gray bars are for travel before and during the pandemic, respectively.
-Considering travel before the pandemic, travel by rickshaw, and bus were relatively common for many respondents (few respondents reported _never_ using these modes).
-The mode most commonly used on a quotidian basis was walk.
-In contrast, respondents reported less frequent travel by car, ridesharing services, CNG auto-rickshaw, motorcycle, and bicycle.
-During the pandemic, we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor.
-The frequency of trip-making by other modes changed more noticeably: the frequency of travel by ridesharing services, rickshaw, CNG auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
-The frequency of walking trips also decreased (fewer respondents report walking almost daily), but the reductions in mobility were not so heavily concentrated at the bottom of the scale.
-
-```{r column-plot-cases-before-after, fig.height=6, echo=FALSE, fig.cap="\\label{fig:column-plot-cases}Number of responses by trip-making frequency class and mode, before and during COVID-19"}
-freq_by_mode_long %>%
-  filter(Mode != "other") %>%
-  ggplot(aes(x = frequency, y = Cases, fill = Period)) +
-  geom_col(color = "black", position = "dodge") + 
-  scale_fill_manual(values = c("before" = "white", "during" = "gray")) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90)) +
-  facet_wrap(. ~ Mode, nrow = 4)
-```
-
-Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of cases in each trip-making frequency class before and during the pandemic.
-If no changes had occurred, all values would be concentrated on the main diagonal of the matrices.
-Values in the lower triangular matrix represent a _loss_ of mobility (lower travel frequency), whereas values in the upper triangular matrix are _gains_ (higher travel frequency).
-The further away a value is from the main diagonal, the greater the loss or gain.
-
-Despite across-the-board losses of mobility, there appears to have been some adaptation that varied by mode of transportation.
-To illustrate, `r bus_xtab[6, 1]` respondents, or `r paste(round(bus_xtab[6, 1] / sum(bus_xtab[6,]) * 100, 2), "%", sep="")` of those who traveled by bus almost daily before, reported never using it during the pandemic.
-In contrast, only `r sum(bus_xtab[1, 2:6])` respondents, or `r paste(round(sum(bus_xtab[1, 2:6]) / sum(bus_xtab[,]) * 100, 2), "%", sep="")` of those who never used buses before started doing so during the pandemic.
-By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so - but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
-
-```{r table-transitions, echo=FALSE}
-rbind(car_xtab, 
-      rideshare_xtab,
-      rickshaw_xtab, 
-      cng_xtab,
-      bus_xtab, 
-      motorcycle_xtab,
-      walk_xtab,
-      bicycle_xtab) %>%
-  kable("latex",
-        booktabs = TRUE,
-        caption = "\\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19") %>%
-  kable_styling(latex_options = c("scale_down")) %>%
-  pack_rows("car", 1, 6) %>%
-  pack_rows("rideshare", 7, 12) %>%
-  pack_rows("rickshaw", 13, 18) %>%
-  pack_rows("cng auto-rickshaw", 19, 24) %>%
-  pack_rows("bus", 25, 30) %>%
-  pack_rows("motorcycle", 31, 36) %>%
-  pack_rows("walk", 37, 42) %>%
-  pack_rows("bicycle", 43, 48)
-```
-
-To more clearly understand the transitions towards different trip-making frequencies, including possible adaptations, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
-
-Figures \ref{fig:circular-plot-1} to \ref{fig:circular-plot-4} present these plots.
-Each of the trip-making frequency sectors on the left hemisphere of the circle represents 100% of responses _before_ the pandemic.
-The links' size is proportional to the probability $p_{ij}$ of transitioning from frequency class $i$ before to the right hemisphere are proportional to the transition probabilities to each frequency _during_ the pandemic.
-There are three transparency levels for the links: solid colors are for $p_{ij}>2/3$, intermediate transparency is for $1/3 < p_{ij} \le 2/3$, and the more transparent links are for $p_{ij}\le 1/3$.
-
-From Figure \ref{fig:circular-plot-1}, we see that the probability of traveling less frequently for those who frequently traveled by car before is high, but their probability of not using this mode at all during the pandemic is quite small.
-The probability of traveling more frequently by car for those who originally never or rarely used this mode remained is low.
-In contrast, we see that the probabilities of never ridesharing during the pandemic are high irrespective of the initial level of use of this mode of transportation.
-
-The probabilities of change in trip frequency by rickshaw and CNG auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}), although the probabilities of being less mobile by CNG auto-rickshaw are greater: between 1/3 and 2/3 of respondents who used this mode almost daily, stopped using it during the pandemic.
-Very rarely there was an increase in mobility by these modes.
-
-After rideshare users, bus users (Figure \ref{fig:circular-plot-3}) were the most likely to stop using it, irrespective of their initial level of use of this mode, while the probabilities of using it _more_ frequently are extremely small.
-
 ```{r reorder-probability-transition-matrices, include=FALSE}
@@
 # one for rows and one for columns
-state_col2 <- c(state_col, rev(state_col))
-names(state_col2) <- c(rownames(bicycle_ptab2), colnames(bicycle_ptab2))
+state_col2 <- c(state_col, 
+                rev(state_col))
+names(state_col2) <- c(rownames(bicycle_ptab2), 
+                       colnames(bicycle_ptab2))
 
-colmat <- rep(state_col2[rownames(bicycle_ptab2)], n_states)
-colmat <- rgb(t(col2rgb(colmat)), maxColorValue = 255)
+colmat <- rep(state_col2[rownames(bicycle_ptab2)], 
+              n_states)
+colmat <- rgb(t(col2rgb(colmat)), 
+              maxColorValue = 255)
 ```
@@
 #Create a grouping variable for the chord diagram, so that sectors are grouped by their state as before and during:
-group = structure(c("Before", "Before", "Before", "Before", "Before", "Before", "During", "During", "During", "During", "During", "During"), names = names(state_col2))
+group = structure(c("Before", 
+                    "Before", 
+                    "Before", 
+                    "Before", 
+                    "Before", 
+                    "Before", 
+                    "During", 
+                    "During", 
+                    "During", 
+                    "During", 
+                    "During", 
+                    "During"),
+                  names = names(state_col2))
 ```
@@
              track.height = 0.2)
-  circos.initialize(factors = factors, xlim = c(0, 1))
+  circos.initialize(factors = factors, 
+                    xlim = c(0, 1))
   circos.track(ylim = c(0, 1), 
@@
 
-  title(paste0("Frequency of travel by ", mode.title, sep = ""),
+  title(paste0("Frequency of travel by ", 
+               mode.title, sep = ""),
         line = nlines,
@@
 
+Research Questions and Hypotheses
+==========================
+
+The spread of the COVID-19 pandemic has led to limitations to movement in many countries and regions, either because of lock-down policies or self-censoring by segments of the public.
+The magnitude of changes in mobility has been studied by recent research, including @DeWeese2020tale and @Molloy2020tracing.
+While the evidence available indicates that although the overall mobility has reduced in much  of the world, the changes were uneven depending on the mode of transportation or the purpose of the trip [see @Lock2020cycling; @Paez2020using].
+Unfortunately, with few exceptions, evidence remains more spotty for developing countries, most of which have large population segments that are less able to absorb losses in mobility [e.g., @Astroza2020mobility; @Huynh2020culture; @Saha2020lockdown].
+
+The objective of this paper is to investigate changes in the trip-making frequency by different modes of transportation during the COVID-19 emergency in Bangladesh.
+Using data from a recent survey that asked respondents to report mobility levels before and during the pandemic, we pose the following questions:
+
+- Was there a reduction of mobility in Bangladesh during COVID-19?
+- And if so, what forms of transportation were more affected?
+
+This paper is a reproducible research document [see @Brunsdon2020opening]; the code and data necessary to reproduce the tables and figures are available in a public repository\footnote{\url{https://github.com/paezha/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh}}
+
+Methods and Data
+============
+
+Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
+The survey was available both in Bengali and English and disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
+In addition to socio-economic and demographic information of the respondents, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
+
+For the purpose of this paper, we use two questions that provide information about trip-making frequency by eight modes of transportation.
+The modes are **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
+Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
+There are $n=800$ responses in the data set.
+
+To describe changes in travel frequency by mode in the transition to the pandemic, we use well-established exploratory data analysis (EDA) techniques.
+
+Findings
+===================
+
+Figure \ref{fig:column-plot-cases} shows the number of responses (out of 800) in each trip-making frequency class by mode of transportation.
+The white bars and gray bars are for travel before and during the pandemic, respectively.
+Considering travel before the pandemic, travel by rickshaw, and bus were relatively common for many respondents (few respondents reported _never_ using these modes).
+The mode most commonly used on a quotidian basis was walk.
+In contrast, respondents reported less frequent travel by car, ridesharing services, CNG auto-rickshaw, motorcycle, and bicycle.
+During the pandemic, we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor.
+
+```{r column-plot-cases-before-after, echo=FALSE, fig.width= 5, fig.cap="\\label{fig:column-plot-cases}Number of responses by trip-making frequency class and mode, before and during COVID-19"}
+freq_by_mode_long %>%
+  filter(Mode != "other") %>%
+  ggplot(aes(x = frequency, y = Cases, fill = Period)) +
+  geom_col(color = "black", position = "dodge") + 
+  scale_fill_manual(values = c("before" = "white", "during" = "gray")) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90)) +
+  facet_wrap(. ~ Mode, nrow = 4)
+```
+
+```{r table-transitions, echo=FALSE}
+rbind(car_xtab, 
+      rideshare_xtab,
+      rickshaw_xtab, 
+      cng_xtab,
+      bus_xtab, 
+      motorcycle_xtab,
+      walk_xtab,
+      bicycle_xtab) %>%
+  kable("latex",
+        booktabs = TRUE,
+        caption = "\\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19") %>%
+  kable_styling(latex_options = c("scale_down", "HOLD_position")) %>%
+  pack_rows("car", 1, 6) %>%
+  pack_rows("rideshare", 7, 12) %>%
+  pack_rows("rickshaw", 13, 18) %>%
+  pack_rows("CNG auto-rickshaw", 19, 24) %>%
+  pack_rows("bus", 25, 30) %>%
+  pack_rows("motorcycle", 31, 36) %>%
+  pack_rows("walk", 37, 42) %>%
+  pack_rows("bicycle", 43, 48)
+```
+
+The frequency of trip-making by other modes changed more noticeably: the frequency of travel by ridesharing services, rickshaw, CNG auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
+The frequency of walking trips also decreased (fewer respondents report walking almost daily), but the reductions in mobility were not so heavily concentrated at the bottom of the scale.
+
+Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of cases in each trip-making frequency class before and during the pandemic.
+If no changes had occurred, all values would be concentrated on the main diagonal of the matrices.
+Values in the lower triangular matrix represent a _loss_ of mobility (lower travel frequency), whereas values in the upper triangular matrix are _gains_ (higher travel frequency).
+The further away a value is from the main diagonal, the greater the loss or gain.
+
+Despite across-the-board losses of mobility, there appears to have been some adaptation that varied by mode of transportation.
+To illustrate, `r bus_xtab[6, 1]` respondents, or `r paste(round(bus_xtab[6, 1] / sum(bus_xtab[6,]) * 100, 2), "%", sep="")` of those who traveled by bus almost daily before, reported never using it during the pandemic.
+In contrast, only `r sum(bus_xtab[1, 2:6])` respondents, or `r paste(round(sum(bus_xtab[1, 2:6]) / sum(bus_xtab[,]) * 100, 2), "%", sep="")` of those who never used buses before started doing so during the pandemic.
+By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so - but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
+
+To more clearly understand the transitions towards different trip-making frequencies, including possible adaptations, we convert the cross-tabulations to probability transition matrices, which we then visualize using circular plots.
+
+Figures \ref{fig:circular-plot-1} to \ref{fig:circular-plot-4} present these plots.
+Each of the trip-making frequency sectors on the left hemisphere of the circle represents 100% of responses _before_ the pandemic.
+The links' size is proportional to the probability $p_{ij}$ of transitioning from frequency class $i$ before to the right hemisphere are proportional to the transition probabilities to each frequency _during_ the pandemic.
+There are three transparency levels for the links: solid colors are for $p_{ij}>2/3$, intermediate transparency is for $1/3 < p_{ij} \le 2/3$, and the more transparent links are for $p_{ij}\le 1/3$.
+
+From Figure \ref{fig:circular-plot-1}, we see that the probability of traveling less by car for those who initially used this mode frequently is high, but their probability of not using this mode at all during the pandemic is quite small.
+In other words, there was a decline in use, but not complete discontinuation of this mode.
+The probability of traveling more frequently by car for those who originally never or rarely used this mode remained is low.
+In contrast, we see that the probabilities of never ridesharing during the pandemic are high irrespective of the initial level of use of this mode of transportation, and the probability of using this mode more are fairly small.
+
 ```{r circular-plots-transition-probabilities-1, echo = FALSE, fig.cap="\\label{fig:circular-plot-1}Transition probabilities in trip frequency from before to during COVID-19: car and rideshare"}
+
 # Rows and columns to arrange several circular plots
@@
 
+The probabilities of change in trip frequency by rickshaw and CNG auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}), although the probabilities of being less mobile by CNG auto-rickshaw are greater: between 1/3 and 2/3 of respondents who used this mode almost daily, stopped using it during the pandemic.
+Very rarely there were increases in mobility by these modes.
+
 ```{r circular-plots-transition-probabilities-2, echo = FALSE, fig.cap="\\label{fig:circular-plot-2}Transition probabilities in trip frequency from before to during COVID-19: rickshaw and cng auto-rickshaw"}
@@
 # CNG
-circle.plot(cng_ptab2, "cng auto-rickshaw", nlines = -3)
+circle.plot(cng_ptab2, "CNG auto-rickshaw", nlines = -3)
 ```
+
+After rideshare, the use of bus had the largest probabilities of discontinuation (Figure \ref{fig:circular-plot-3}).
+We see that even respondents who used this mode almost daily before the pandemic had close to 66% chances of never using it during the contingency.
+Similarly large losses of mobility by bus were observed for travelers who used this mode less frequently before the pandemic.
+These losses were not offset to any appreciable degree by the probability of some users turning to this mode during the pandemic.
 
@@
 
+Trip-making by motorcycle (Figure \ref{fig:circular-plot-3}) also declined for all but the most frequent users pre-COVID-19, and the probability of adopting this mode during COVID-19 was almost null.
+Contrast this to the case of walking (Figure \ref{fig:circular-plot-4}), where a sizable number of respondents started walking during the pandemic, even if only rarely.
+Close to 50% of respondents who report walking almost daily during the pandemic walked less frequently before the contingency.
+Bicycle (Figure \ref{fig:circular-plot-4}), like motorcycle, saw relatively fewer respondents adopting it during the pandemic, and many travelers who did use the mode rarely, once a week, or even 2-3 times per week stopped cycling during the pandemic.
+
 ```{r circular-plots-transition-probabilities-4, echo = FALSE, fig.cap="\\label{fig:circular-plot-4}Transition probabilities in trip frequency from before to during COVID-19: bicycle and walk"}
@@
 
+The findings reported here provide important changes in mobility in a developing country during COVID-19.
+In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (rideshare, bus, rickshaw, and CNG auto-rickshaw) were more affected.
+Travel by car, motorcycle, and bicycle were somewhat less affected, and there is evidence of some adoption of walking during the pandemic.
+
 References {#references .unnumbered}

### 2020-10-06 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.tex
v1 Paper submitted to Transport Findings

@@
 % Pandoc header
+\usepackage{float} \floatplacement{figure}{H}
 \usepackage{booktabs}
@@
   The COVID-19 pandemic has had a profound impact on mobility in every
-  country and region around the world. Recent research provides
-  illuminates the nature and magnitude of the changes in mobility, but the
-  evidence is still scant in developing countries. The objective of this
-  paper is to present an exploratory analysis of the changes in the
-  frequency of trip-making by mode during the COVID-19 emergency in
-  Bangladesh. The results confirm an overall loss of mobility in the form
-  of reduced trip-making frequency by all modes, but the changes are not
-  uniform across modes.
+  country and region around the world. Recent helps to illuminate some of
+  the dimensions of change - however, the evidence is still scant in
+  developing countries. The objective of this paper is to present an
+  exploratory analysis of the changes in the frequency of trip-making by
+  mode during the COVID-19 emergency in Bangladesh. The results confirm an
+  overall loss of mobility in the form of reduced trip-making frequency by
+  all modes, but the changes are not uniform across modes. In particular,
+  there was a greater loss of mobility for bus, rickshaw, and CNG
+  auto-rickshaw.
   \end{abstract}
@@
 (2020) and Molloy et al. (2020). While the evidence available indicates
-that overall there was a reduction in mobility in much of the world, the
+that although the overall mobility has reduced in much of the world, the
 changes were uneven depending on the mode of transportation or the
-purpose of the trip (see Lock, 2020; Paez, 2020). Alas, with few
-exceptions evidence remains more spotty for developing countries, most
-of which have large populations segments that are less able to absorb
-losses in mobility (e.g., Astroza et al., 2020; Huynh, 2020; Saha et
-al., 2020).
+purpose of the trip (see Lock, 2020; Paez, 2020). Unfortunately, with
+few exceptions, evidence remains more spotty for developing countries,
+most of which have large population segments that are less able to
+absorb losses in mobility (e.g., Astroza et al., 2020; Huynh, 2020; Saha
+et al., 2020).
 
@@
 This paper is a reproducible research document (see Brunsdon and Comber,
-2020); the code and data necessary to replicate the tables and figures
-are available in a public repository\footnote{\url{place.holder}}
+2020); the code and data necessary to reproduce the tables and figures
+are available in a public
+repository\footnote{\url{https://github.com/paezha/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh}}
 
@@
 
-Data used for this paper come from a survey conducted during
-\textbf{DATES} in \textbf{regions/cities} in Bangladesh. \textbf{BRIEFLY
-ADD DETAILS ABOUT SURVEY}.
+Data used for this paper come from an online survey, ``Exploring the
+potential of travel mode change behavior in the post-lock-down and
+post-pandemic (COVID-19) period'' conducted during July - August 2020 in
+Bangladesh. The survey was available both in Bengali and English and
+disseminated through various electronic means along with a widespread
+social media campaign and promotion in various social media sites and
+groups. In addition to socio-economic and demographic information of the
+respondents, the survey covered different aspects of travel such as
+travel behavior before and during COVID-19, knowledge related to
+COVID-19, and opinions and perceptions regarding travel behavior after
+lock-down and post-pandemic situation.
 
-The survey asked respondents to self-report their trip-making frequency
-by eight modes of transportation, namely \textbf{car},
-\textbf{ridesharing} (e.g., Uber, Pathao), \textbf{rickshaw},
-\textbf{cng auto-rickshaw} (a rickshaw-like vehicle powered by
-compressed natural gas), \textbf{bus}, motorcycle/scooter (hereafter
-just \textbf{motorcycle}), \textbf{walking}, and \textbf{bicycle} (there
-was an additional catch-all category \textbf{other} which we ignore
-here). Participants in the survey used the following levels to report
-their frequency of traveling by each mode both before and during
-COVID-19: \emph{Never}, \emph{Rarely}, \emph{Once a week}, \emph{2-3
-trips per week}, \emph{4-5 trips per week}, \emph{Almost daily}. There
-are \(n=800\) responses in the data set.
+For the purpose of this paper, we use two questions that provide
+information about trip-making frequency by eight modes of
+transportation. The modes are \textbf{car}, \textbf{ridesharing} (e.g.,
+Uber, Pathao), \textbf{rickshaw}, \textbf{CNG auto-rickshaw} (a
+rickshaw-like vehicle powered by compressed natural gas), \textbf{bus},
+motorcycle/scooter (hereafter just \textbf{motorcycle}),
+\textbf{walking}, and \textbf{bicycle} (there was an additional
+catch-all category \textbf{other} which we ignore here). Participants in
+the survey used the following levels to report their frequency of
+traveling by each mode both before and during COVID-19: \emph{Never},
+\emph{Rarely}, \emph{Once a week}, \emph{2-3 trips per week}, \emph{4-5
+trips per week}, \emph{Almost daily}. There are \(n=800\) responses in
+the data set.
 
-To describe changes in the frequency of travel by mode in the transition
-to the pandemic, we use well-established exploratory data analysis (EDA)
+To describe changes in travel frequency by mode in the transition to the
+pandemic, we use well-established exploratory data analysis (EDA)
 techniques.
@@
 respectively. Considering travel before the pandemic, travel by
-rickshaw, bus were relatively common for many respondents (few
+rickshaw, and bus were relatively common for many respondents (few
 respondents reported \emph{never} using these modes). The mode most
 commonly used on a quotidian basis was walk. In contrast, respondents
-reported less frequent travel by car, rideshare services, cng
-auto-rickshaw, motorcycle, and bicycle. During the pandemic we see that
+reported less frequent travel by car, ridesharing services, CNG
+auto-rickshaw, motorcycle, and bicycle. During the pandemic, we see that
 while there were reductions in mobility by car, motorcycle, and bicycle
 (with more respondents reporting never traveling by these modes), the
-changes were relatively minor. The frequency of trip-making by other
-modes changed more noticeably: the frequency of travel by rideshare
-services, rickshaw, cng auto-rickshaw, and bus collapsed, with vastly
-more respondents reporting never using these modes during the pandemic
-than before. The frequency of walking trips also decreased (fewer
-respondents report walking almost daily), but the reductions in mobility
-were not so heavily concentrated at the bottom of the scale.
+changes were relatively minor.
 
@@
 
-Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of
-cases in each trip-making frequency class before and during the
-pandemic. If no changes had occurred, all values would be concentrated
-on the main diagonal of the matrices. Values in the lower triangular
-matrix represent a \emph{loss} of mobility (lower travel frequency),
-whereas values in the upper triangular matrix are \emph{gains} (higher
-travel frequency). The further away a value is from the main diagonal,
-the greater the loss or gain.
-
-Despite across-the-board losses of mobility, there appears to have been
-some adaptation that varied by mode of transportation. To illustrate,
-103 respondents, or 65.61\% of those who traveled by bus almost daily
-before, reported never using it during the pandemic. In contrast, only
-12 respondents, or 1.5\% of those who never used buses before started
-doing so during the pandemic. By way of comparison, 24.14\% of
-respondents who cycled almost daily before the pandemic stopped doing so
-- but 4.88\% who never cycled before started doing so during the
-pandemic.
-
-\begin{table}
+\begin{table}[H]
 
@@
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{cng auto-rickshaw}}\\
+\multicolumn{7}{l}{\textbf{CNG auto-rickshaw}}\\
 \hspace{1em}Never & 176 & 14 & 3 & 1 & 0 & 0\\
@@
 
-To more easily understand the transitions towards different trip-making
-frequencies, including possible adaptations by mode, we convert the
+The frequency of trip-making by other modes changed more noticeably: the
+frequency of travel by ridesharing services, rickshaw, CNG
+auto-rickshaw, and bus collapsed, with vastly more respondents reporting
+never using these modes during the pandemic than before. The frequency
+of walking trips also decreased (fewer respondents report walking almost
+daily), but the reductions in mobility were not so heavily concentrated
+at the bottom of the scale.
+
+Table \ref{tab:cross-tabulation} is a cross-tabulation of the number of
+cases in each trip-making frequency class before and during the
+pandemic. If no changes had occurred, all values would be concentrated
+on the main diagonal of the matrices. Values in the lower triangular
+matrix represent a \emph{loss} of mobility (lower travel frequency),
+whereas values in the upper triangular matrix are \emph{gains} (higher
+travel frequency). The further away a value is from the main diagonal,
+the greater the loss or gain.
+
+Despite across-the-board losses of mobility, there appears to have been
+some adaptation that varied by mode of transportation. To illustrate,
+103 respondents, or 65.61\% of those who traveled by bus almost daily
+before, reported never using it during the pandemic. In contrast, only
+12 respondents, or 1.5\% of those who never used buses before started
+doing so during the pandemic. By way of comparison, 24.14\% of
+respondents who cycled almost daily before the pandemic stopped doing so
+- but 4.88\% who never cycled before started doing so during the
+pandemic.
+
+To more clearly understand the transitions towards different trip-making
+frequencies, including possible adaptations, we convert the
 cross-tabulations to probability transition matrices, which we then
@@
 these plots. Each of the trip-making frequency sectors on the left
-hemisphere of the circle represent 100\% of responses \emph{before} the
-pandemic. The size of the links is proportional to the probability
-\(p_{ij}\) of transitioning from frequency class \(i\) before to the
-right hemisphere are proportional to the transition probabilities to
-each frequency \emph{during} the pandemic. There are three transparency
+hemisphere of the circle represents 100\% of responses \emph{before} the
+pandemic. The links' size is proportional to the probability \(p_{ij}\)
+of transitioning from frequency class \(i\) before to the right
+hemisphere are proportional to the transition probabilities to each
+frequency \emph{during} the pandemic. There are three transparency
 levels for the links: solid colors are for \(p_{ij}>2/3\), intermediate
@@
 
-From Figure \ref{fig:circular-plot-1} we see that the probability of
-traveling less frequently for those who frequently traveled by car
-before is high, but their probability of not using this mode at all
-during the pandemic is quite small. The probability of traveling more
-frequently by car for those who originally never or rarely used this
-mode remained is low. In contrast, we see that the probabilities of
+From Figure \ref{fig:circular-plot-1}, we see that the probability of
+traveling less by car for those who initially used this mode frequently
+is high, but their probability of not using this mode at all during the
+pandemic is quite small. In other words, there was a decline in use, but
+not complete discontinuation of this mode. The probability of traveling
+more frequently by car for those who originally never or rarely used
+this mode remained is low. In contrast, we see that the probabilities of
 never ridesharing during the pandemic are high irrespective of the
-initial level of use of this mode of transportation.
-
-The probabilities of change in trip frequency by rickshaw and and cng
-auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}),
-although the probabilities of being less mobile by cng auto-rickshaw are
-greater: between 1/3 and 2/3 of respondents who used this mode almost
-daily, stopped using it during the pandemic. Very rarely there was an
-increase in mobility by these modes.
-
-After ridesharers, bus users (Figure \ref{fig:circular-plot-3}) were the
-most likely to stop using it, irrespective of their initial level of use
-of this mode, while the probabilities of using it \emph{more} frequently
-are extremely small.
+initial level of use of this mode of transportation, and the probability
+of using this mode more are fairly small.
 
@@
 \end{figure}
+
+The probabilities of change in trip frequency by rickshaw and CNG
+auto-rickshaw are similar (see Figure \ref{fig:circular-plot-2}),
+although the probabilities of being less mobile by CNG auto-rickshaw are
+greater: between 1/3 and 2/3 of respondents who used this mode almost
+daily, stopped using it during the pandemic. Very rarely there were
+increases in mobility by these modes.
 
@@
 
+After rideshare, the use of bus had the largest probabilities of
+discontinuation (Figure \ref{fig:circular-plot-3}). We see that even
+respondents who used this mode almost daily before the pandemic had
+close to 66\% chances of never using it during the contingency.
+Similarly large losses of mobility by bus were observed for travelers
+who used this mode less frequently before the pandemic. These losses
+were not offset to any appreciable degree by the probability of some
+users turning to this mode during the pandemic.
+
 \begin{figure}
@@
 
+Trip-making by motorcycle (Figure \ref{fig:circular-plot-3}) also
+declined for all but the most frequent users pre-COVID-19, and the
+probability of adopting this mode during COVID-19 was almost null.
+Contrast this to the case of walking (Figure \ref{fig:circular-plot-4}),
+where a sizable number of respondents started walking during the
+pandemic, even if only rarely. Close to 50\% of respondents who report
+walking almost daily during the pandemic walked less frequently before
+the contingency. Bicycle (Figure \ref{fig:circular-plot-4}), like
+motorcycle, saw relatively fewer respondents adopting it during the
+pandemic, and many travelers who did use the mode rarely, once a week,
+or even 2-3 times per week stopped cycling during the pandemic.
+
 \begin{figure}
@@
 \end{figure}
+
+The findings reported here provide important changes in mobility in a
+developing country during COVID-19. In general there was a loss of
+mobility across-the-board, but modes that require interaction with
+strangers (rideshare, bus, rickshaw, and CNG auto-rickshaw) were more
+affected. Travel by car, motorcycle, and bicycle were somewhat less
+affected, and there is evidence of some adoption of walking during the
+pandemic.
 

### 2020-10-06 README.Rmd
v1 Created README.Rmd

--- 2020-10-06 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
@@
 ---
-title: An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
-author:
-  - name: Shaila Jamal
-    email: jamals16@mcmaster.ca
-    affiliation: McMaster University
-  - name: Antonio Paez
-    email: paezha@mcmaster.ca
-    affiliation: McMaster University
-    footnote: 1
-address:
-  - code: McMaster University
-    address: School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
+output: github_document
+always_allow_html: true
+bibliography: Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/References.bib
+---
+
+<!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+```
+
+# Changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
+
+<!-- badges: start -->
+<!-- badges: end -->
+
+Shaila Jamal (jamals16@mcmaster.ca)  
+Antonio Paez (paezha@mcmaster.ca)  
+School of Geography and Earth Science  
+McMaster University  
+1280 Main Street West, Hamilton, ON L8S 4K1  
+Canada
+
+Paper submitted for peer review to Transport Findings
+
+## Abstract
+
 The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
@@
 The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
-In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw.
+In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw, and some adoption of walking during the pandemic.
 
-journal: "Transport Findings"
-date: "`r Sys.Date()`"
-bibliography: References.bib
-#linenumbers: true
-#numbersections: true
-csl: elsevier-harvard.csl
-output: rticles::elsevier_article
-always_allow_html: true
+## Keywords
 
-header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
----
+- COVID-19
+- Mobility
+- Trip-making frequency
+- Modes of transportation
+- Transition probabilities
+- Circular plots
 
@@
 
-```{r read-data, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Read data file, add unique respondent identifier:
-mcb <- read_csv("Combined-data-final-with-serial-number.csv") %>% 
-  rename(id = `Serial no.`,
-         yob = `24.	Please enter your year of birth (e.g. 1990, 1999)`)
-```
-
-```{r convert-to-factors-and-clean, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Extract relevant variables and convert to factors as appropriate. A small number of records (4) are deleted because they contain nonsensical year of birth responses 
-freq_by_mode <- mcb %>%
-  mutate(id = c(1:nrow(mcb)),
-         age = 2020 - yob) %>%
-  filter(age > 0 & age < 100) %>%
-  transmute(id,
-            freq_car_before = factor(.[[20]]),
-            freq_bus_before = factor(.[[21]]),
-            freq_rickshaw_before = factor(.[[22]]),
-            freq_cng_before = factor(.[[23]]),
-            freq_walk_before = factor(.[[24]]),
-            freq_motorcycle_before = factor(.[[25]]),
-            freq_bicycle_before = factor(.[[26]]),
-            freq_rideshare_before = factor(.[[27]]),
-            freq_other_before = factor(.[[28]]),
-            freq_car_during = factor(.[[29]]),
-            freq_bus_during = factor(.[[30]]),
-            freq_rickshaw_during = factor(.[[31]]),
-            freq_cng_during = factor(.[[32]]),
-            freq_walk_during = factor(.[[33]]),
-            freq_motorcycle_during = factor(.[[34]]),
-            freq_bicycle_during = factor(.[[35]]),
-            freq_rideshare_during = factor(.[[36]]),
-            freq_other_during = factor(.[[37]]))
-```
-
-```{r frequency-table-pivot-long, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Pivot table to long: each row is a respondent-mode-period response
-freq_by_mode_long <- freq_by_mode %>%
-  pivot_longer(cols = starts_with("freq_"),
-               names_to = "mode_period",
-               names_prefix = "freq_",
-               values_to = "frequency") %>%
-  separate(mode_period, c("Mode", "Period")) %>%
-  transmute(id,
-            Period = factor(Period,
-                            levels = c("before", "during")), 
-            Mode = factor(Mode,
-                          levels = c("car",
-                                     "motorcycle",
-                                     "rideshare",
-                                     "rickshaw",
-                                     "cng",
-                                     "bus",
-                                     "walk",
-                                     "bicycle",
-                                     "other")), 
-            frequency = factor(frequency, 
-                               levels = c("Never",
-                                          "Rarely",
-                                          "At least once a week",
-                                          "2-3 times a week",
-                                          "4-5 times a week",
-                                          "Almost everyday"),
-                               labels = c("Never",
-                                          "Rarely",
-                                          "Once a week",
-                                          "2-3 per week",
-                                          "4-5 per week",
-                                          "Almost daily"),
-                               ordered = TRUE))
-```
-
-```{r frequency-table-pivot-wide, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Pivot the table to wide: each row is a respondent
-freq_by_mode <- freq_by_mode_long %>%
-  pivot_wider(names_from = c("Mode", "Period"),
-              names_sep = "_",
-              names_prefix = "freq_",
-              values_from = "frequency")
-```
-
-```{r summarize-frequency, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Summarize cases by period and mode
-freq_by_mode_long <- freq_by_mode_long %>% 
-  group_by(Period, Mode, frequency) %>%
-  summarize(Cases = n())
-```
-
-```{r probability-transitions, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Create cross-tabulations and tables of proportions by mode. The cross-tabulation is of the number of cases by period (before COVID-19 and during):
-
-# CAR
-car_xtab <- xtabs(~ freq_car_before + freq_car_during, 
-                  data = freq_by_mode)
-
-car_ptab <- car_xtab %>%
-  prop.table(1)
-
-# MOTORCYCLE
-motorcycle_xtab <- xtabs(~ freq_motorcycle_before + freq_motorcycle_during, 
-                         data = freq_by_mode)
-
-motorcycle_ptab <- motorcycle_xtab %>%
-  prop.table(1)
-
-# RIDESHARE
-rideshare_xtab <- xtabs(~ freq_rideshare_before + freq_rideshare_during, 
-                        data = freq_by_mode)
-
-rideshare_ptab <- rideshare_xtab %>%
-  prop.table(1)
-
-# RICKSHAW
-rickshaw_xtab <- xtabs(~ freq_rickshaw_before + freq_rickshaw_during, 
-                       data = freq_by_mode)
-
-rickshaw_ptab <- rickshaw_xtab %>%
-  prop.table(1)
-
-# cng
-cng_xtab <- xtabs(~ freq_cng_before + freq_cng_during, 
-                  data = freq_by_mode)
-
-cng_ptab <- cng_xtab %>%
-  prop.table(1)
-
-# BUS
-bus_xtab <- xtabs(~ freq_bus_before + freq_bus_during, 
-                  data = freq_by_mode)
-
-bus_ptab <- bus_xtab %>%
-  prop.table(1)
-
-# WALK
-walk_xtab <- xtabs(~ freq_walk_before + freq_walk_during, 
-                   data = freq_by_mode)
-
-walk_ptab <- walk_xtab %>%
-  prop.table(1)
-
-# BICYCLE
-bicycle_xtab <- xtabs(~ freq_bicycle_before + freq_bicycle_during, 
-                      data = freq_by_mode)
-
-bicycle_ptab <- bicycle_xtab %>%
-  prop.table(1)
-
-# OTHER
-other_xtab <- xtabs(~ freq_other_before + freq_other_during, 
-                    data = freq_by_mode)
-
-other_ptab <- other_xtab %>%
-  prop.table(1)
-```
-
-```{r save-data, include=FALSE, eval=FALSE}
-# NOT RUN - THE DATA FILES ARE IN PRIVATE DIRECTORY. INCLUDED ONLY FOR TRANSPARENCY OF DATA PREPARATION
-
-# Save data objects
-save(freq_by_mode_long, 
-     car_xtab, 
-     car_ptab, 
-     motorcycle_xtab, 
-     motorcycle_ptab, 
-     rideshare_xtab, 
-     rideshare_ptab, 
-     rickshaw_xtab, 
-     rickshaw_ptab, 
-     cng_xtab, 
-     cng_ptab, 
-     bus_xtab, 
-     bus_ptab, 
-     walk_xtab, 
-     walk_ptab, 
-     bicycle_xtab, 
-     bicycle_ptab, 
-     other_xtab, 
-     other_ptab,
-     file = "Trip-Frequency-Bangladesh.RData")
-```
-
 ```{r load-data-file, include=FALSE}
-load("Trip-Frequency-Bangladesh.RData")
+load("Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Trip-Frequency-Bangladesh.RData")
 ```
@@
 In contrast, only `r sum(bus_xtab[1, 2:6])` respondents, or `r paste(round(sum(bus_xtab[1, 2:6]) / sum(bus_xtab[,]) * 100, 2), "%", sep="")` of those who never used buses before started doing so during the pandemic.
-By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so - but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
+By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so, but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
 

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Response-to-Reviewers-v2.Rmd
v2 Started working on revisions after reviews

---
title: "Response to Reviewers"
author: "Shaila Jamal and Antonio Paez"
date: "5/11/2020"
output: pdf_document
---

We wish to thank two anonymous reviewers for their comments and suggestions for revisions. In this letter we respond to those comments and explain how the paper was revised in response.

\newpage

# Reviewer 1

The paper presents an exploratory analysis of online survey results by 800 respondents on changes to their trip making behaviours by mode during the COVD pandemic. It is a well written paper, and I recommend publication. Some minor edits and adds are suggested below.

\textcolor{blue}{Thank you for your positive assessment and comments for improving the paper.}

- Figure 1: Please lengthen the y-axes for all figures, at the moment, it is a bit hard to read results off, because the axes are squashed.

\textcolor{blue}{Done. Hopefully the aspect of the figure will make it easier to read the results.}

- Table 1: Please mark whether the rows are before and columns are during in the cross tabulation. It would be a good idea to use something like a conditional colouring so that each cell value is coloured by the scale of the value, so that the patterns discussed come out visually clear. At the moment, one has to read the numbers carefully.

\textcolor{blue}{Words go here.}

- Otherwise, overall the paper is well written and the results are clearly discussed.

- Bangladesh has very high density urban areas, it would be good to include a discussion on whether the online survey responses indicated which cities the respondents are from - what was the geography of the answers? While 800 is a decent number, the volume is still a small number when compared as a sample against full populations, further, different cities have different transit arrangements, and since the shared and transit modes seem most affected - a discussion of geography somewhere (which principal cities and their attributes) will enrich the discussion, and also as mentioned a discussion of the limitations of the sample size being still relatively small.

\textcolor{blue}{Words go here.}

\newpage

# Reviewer 2

The level of reliability of the conclusions critically depends on the reliability of the sample, which is not analysed at all, i.e., there is no analysis of the level of representativeness of the sample relative to the Bangladesh population. A social media sample likely overrepresents high incomes, young age, etc. This must be scrutinized.

This is relevant because in the paper, from the title to the summary in the last paragraph, the authors claim they represent mobility changes in a developing country: Bangladesh, which has 160 million people (very small sample=800)

Other comments:

- No information of what contagion containment measures, if any, were in place during the time of the survey (e.g., lockdown, school closures)

- Table 1: columns and rows need to indicate which one is pre and during COVID, it has to be added to the table.

- Uber is not ridesharing, it is ridesourcing or ride-hailing.

\textcolor{blue}{This has been changed throughout the paper.}

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
v2 Started working on revisions after reviews

@@
 ---
-title: An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
+title: Changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
 author:
@@
 The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
-In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw.
+In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw, and some adoption of walking during the pandemic.
 
@@
  \usepackage{float}
+ \usepackage{xcolor}
  \floatplacement{figure}{H}
+
 ---
@@
                                      "bicycle",
+                                     "other"),
+                          labels = c("car",
+                                     "motorcycle",
+                                     "ride-hailing",
+                                     "rickshaw",
+                                     "cng",
+                                     "bus",
+                                     "walk",
+                                     "bicycle",
                                      "other")), 
@@
 
-# RIDESHARE
+# RIDE-HAILING
 rideshare_xtab <- xtabs(~ freq_rideshare_before + freq_rideshare_during, 
@@
 
-# RIDESHARE
+# RIDE-HAILING
 rownames(rideshare_ptab2) = paste0("B_", all_states)
@@
 For the purpose of this paper, we use two questions that provide information about trip-making frequency by eight modes of transportation.
-The modes are **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
+The modes are **car**, **ride-hailing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
 Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
@@
 The mode most commonly used on a quotidian basis was walk.
-In contrast, respondents reported less frequent travel by car, ridesharing services, CNG auto-rickshaw, motorcycle, and bicycle.
+In contrast, respondents reported less frequent travel by car, ride-hailing services, CNG auto-rickshaw, motorcycle, and bicycle.
 During the pandemic, we see that while there were reductions in mobility by car, motorcycle, and bicycle (with more respondents reporting never traveling by these modes), the changes were relatively minor.
 
-```{r column-plot-cases-before-after, echo=FALSE, fig.width= 5, fig.cap="\\label{fig:column-plot-cases}Number of responses by trip-making frequency class and mode, before and during COVID-19"}
+```{r column-plot-cases-before-after, echo=FALSE, fig.height= 7, fig.cap="\\label{fig:column-plot-cases}Number of responses by trip-making frequency class and mode, before and during COVID-19"}
 freq_by_mode_long %>%
@@
   facet_wrap(. ~ Mode, nrow = 4)
+
+# Uncomment to save to jpg files for the publisher
+#ggsave(filename = "Figure-1.jpeg", device = "jpeg", height = 9, width = 6.5, units = "in")
 ```
@@
       bicycle_xtab) %>% 
+  t() %>%
+  data.frame() %>% 
+  mutate(across(.cols = everything(), function(x) {
+    cell_spec(x, "latex", bold = TRUE, color = "white", background = spec_color(x, end = 0.8, alpha = 0.3, option = "plasma", direction = -1))
+    })) %>%
+  t() %>%
+  data.frame() %>%
+  rename(Never = X1, Rarely = X2, `Once a week` = X3, `2-3 times per week` = X4, `4-5 times per week` = X5, `Almost daily` = X6) %>%
+  rownames_to_column(var = "Before") %>%
+  mutate(Before = rep(c("Never", "Rarely", "Once a week", "2-3 times per week", "4-5 times per week", "Almost daily"), 8)) %>%
   kable("latex",
         booktabs = TRUE,
+        escape = FALSE,
         caption = "\\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19") %>%
   kable_styling(latex_options = c("scale_down", "HOLD_position")) %>%
+  add_header_above(c(" " = 1, "During" = 6)) %>%
   pack_rows("car", 1, 6) %>%
-  pack_rows("rideshare", 7, 12) %>%
+  pack_rows("ride-hailing", 7, 12) %>%
   pack_rows("rickshaw", 13, 18) %>%
@@
   pack_rows("walk", 37, 42) %>%
-  pack_rows("bicycle", 43, 48)
+  pack_rows("bicycle", 43, 48)  %>%
+  footnote(general = "Darker cell colors indicate higher cell values in a row.")
 ```
 
-The frequency of trip-making by other modes changed more noticeably: the frequency of travel by ridesharing services, rickshaw, CNG auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
+The frequency of trip-making by other modes changed more noticeably: the frequency of travel by ride-hailing services, rickshaw, CNG auto-rickshaw, and bus collapsed, with vastly more respondents reporting never using these modes during the pandemic than before.
 The frequency of walking trips also decreased (fewer respondents report walking almost daily), but the reductions in mobility were not so heavily concentrated at the bottom of the scale.
@@
 The probability of traveling more frequently by car for those who originally never or rarely used this mode remained is low.
-In contrast, we see that the probabilities of never ridesharing during the pandemic are high irrespective of the initial level of use of this mode of transportation, and the probability of using this mode more are fairly small.
+In contrast, we see that the probabilities of never ride-hailing during the pandemic are high irrespective of the initial level of use of this mode of transportation, and the probability of using this mode more are fairly small.
 
-```{r circular-plots-transition-probabilities-1, echo = FALSE, fig.cap="\\label{fig:circular-plot-1}Transition probabilities in trip frequency from before to during COVID-19: car and rideshare"}
+```{r circular-plots-transition-probabilities-1, echo = FALSE, fig.cap="\\label{fig:circular-plot-1}Transition probabilities in trip frequency from before to during COVID-19: car and ride-hailing"}
+# Uncomment to save to jpg files for the publisher
+#jpeg("Figure-2.jpg", width = 6.5, height = 4.5, units = "in", res = 600)
 
@@
 
-# RIDESHARE
-circle.plot(rideshare_ptab2, "rideshare", nlines = -3)
+# RIDE-HAIL
+circle.plot(rideshare_ptab2, "ride-hailing", nlines = -3)
+
+# Uncomment to save to jpg files for the publisher
+#dev.off()
 ```
@@
 ```{r circular-plots-transition-probabilities-2, echo = FALSE, fig.cap="\\label{fig:circular-plot-2}Transition probabilities in trip frequency from before to during COVID-19: rickshaw and cng auto-rickshaw"}
+# Uncomment to save to jpg files for the publisher
+#jpeg("Figure-3.jpg", width = 6.5, height = 4.5, units = "in", res = 600)
+
 # Rows and columns to arrange several circular plots
@@
 circle.plot(cng_ptab2, "CNG auto-rickshaw", nlines = -3)
+
+# Uncomment to save to jpg files for the publisher
+#dev.off()
 ```
 
-After rideshare, the use of bus had the largest probabilities of discontinuation (Figure \ref{fig:circular-plot-3}).
+After ride-hailing, the use of bus had the largest probabilities of discontinuation (Figure \ref{fig:circular-plot-3}).
 We see that even respondents who used this mode almost daily before the pandemic had close to 66% chances of never using it during the contingency.
@@
 ```{r circular-plots-transition-probabilities-3, echo = FALSE, fig.cap="\\label{fig:circular-plot-3}Transition probabilities in trip frequency from before to during COVID-19: motorcycle and bus"}
+# Uncomment to save to jpg files for the publisher
+#jpeg("Figure-4.jpg", width = 6.5, height = 4.5, units = "in", res = 600)
+
 # Rows and columns to arrange several circular plots
@@
 circle.plot(motorcycle_ptab2, "motorcycle", nlines = -3)
+
+# Uncomment to save to jpg files for the publisher
+#dev.off()
 ```
@@
 ```{r circular-plots-transition-probabilities-4, echo = FALSE, fig.cap="\\label{fig:circular-plot-4}Transition probabilities in trip frequency from before to during COVID-19: bicycle and walk"}
+# Uncomment to save to jpg files for the publisher
+#jpeg("Figure-5.jpg", width = 6.5, height = 4.5, units = "in", res = 600)
+
 # Rows and columns to arrange several circular plots
@@
 circle.plot(bicycle_ptab2, "bicycle", nlines = -3)
+
+# Uncomment to save to jpg files for the publisher
+#dev.off()
 ```
@@
 The findings reported here provide important changes in mobility in a developing country during COVID-19.
-In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (rideshare, bus, rickshaw, and CNG auto-rickshaw) were more affected.
+In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more affected.
 Travel by car, motorcycle, and bicycle were somewhat less affected, and there is evidence of some adoption of walking during the pandemic.

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.tex
v2 Started working on revisions after reviews

@@
             pdfauthor={},
-            pdftitle={An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh},
+            pdftitle={Changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh},
             colorlinks=false,
@@
 % Pandoc header
-\usepackage{float} \floatplacement{figure}{H}
+\usepackage{float} \usepackage{xcolor} \floatplacement{figure}{H}
 \usepackage{booktabs}
@@
 
-  \title{An exploratory analysis of changes in trip-making frequency by mode
-during the COVID-19 emergency in Bangladesh}
+  \title{Changes in trip-making frequency by mode during the COVID-19 emergency
+in Bangladesh}
     \author[McMaster University]{Shaila Jamal}
@@
   there was a greater loss of mobility for bus, rickshaw, and CNG
-  auto-rickshaw.
+  auto-rickshaw, and some adoption of walking during the pandemic.
   \end{abstract}
@@
 information about trip-making frequency by eight modes of
-transportation. The modes are \textbf{car}, \textbf{ridesharing} (e.g.,
+transportation. The modes are \textbf{car}, \textbf{ride-hailing} (e.g.,
 Uber, Pathao), \textbf{rickshaw}, \textbf{CNG auto-rickshaw} (a
@@
 commonly used on a quotidian basis was walk. In contrast, respondents
-reported less frequent travel by car, ridesharing services, CNG
+reported less frequent travel by car, ride-hailing services, CNG
 auto-rickshaw, motorcycle, and bicycle. During the pandemic, we see that
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrrrrrr}
+\begin{tabular}[t]{lllllll}
 \toprule
-  & Never & Rarely & Once a week & 2-3 per week & 4-5 per week & Almost daily\\
+\multicolumn{1}{c}{ } & \multicolumn{6}{c}{During} \\
+\cmidrule(l{3pt}r{3pt}){2-7}
+Before & Never & Rarely & Once a week & 2-3 times per week & 4-5 times per week & Almost daily\\
 \midrule
@@
 \multicolumn{7}{l}{\textbf{car}}\\
-\hspace{1em}Never & 267 & 20 & 5 & 1 & 1 & 2\\
-\hspace{1em}Rarely & 142 & 99 & 21 & 12 & 9 & 3\\
-\hspace{1em}Once a week & 11 & 23 & 19 & 7 & 2 & 3\\
-\hspace{1em}2-3 per week & 10 & 23 & 7 & 6 & 8 & 2\\
-\hspace{1em}4-5 per week & 4 & 9 & 5 & 6 & 12 & 4\\
-\hspace{1em}Almost daily & 6 & 15 & 6 & 10 & 0 & 20\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{267}}} & \cellcolor[HTML]{F79242}{\textcolor{white}{\textbf{20}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA537}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{142}}} & \cellcolor[HTML]{7D03A8}{\textcolor{white}{\textbf{99}}} & \cellcolor[HTML]{F1824C}{\textcolor{white}{\textbf{21}}} & \cellcolor[HTML]{F79442}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{F99A3E}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{C13B82}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{23}}} & \cellcolor[HTML]{5801A4}{\textcolor{white}{\textbf{19}}} & \cellcolor[HTML]{E26660}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{F9993E}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{CA457A}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{23}}} & \cellcolor[HTML]{E26660}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{E97257}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{8}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{910EA3}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{F1844C}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{E16462}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{6A00A8}{\textcolor{white}{\textbf{15}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{B12A90}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{20}}}\\
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{rideshare}}\\
-\hspace{1em}Never & 254 & 13 & 6 & 4 & 0 & 0\\
-\hspace{1em}Rarely & 192 & 44 & 12 & 1 & 1 & 1\\
-\hspace{1em}Once a week & 60 & 20 & 31 & 3 & 1 & 0\\
-\hspace{1em}2-3 per week & 49 & 12 & 9 & 7 & 2 & 0\\
-\hspace{1em}4-5 per week & 28 & 10 & 3 & 5 & 6 & 2\\
-\hspace{1em}Almost daily & 14 & 3 & 4 & 1 & 0 & 2\\
+\multicolumn{7}{l}{\textbf{ride-hailing}}\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{254}}} & \cellcolor[HTML]{F9983E}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{FB9F3A}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{192}}} & \cellcolor[HTML]{E56A5D}{\textcolor{white}{\textbf{44}}} & \cellcolor[HTML]{F89540}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{60}}} & \cellcolor[HTML]{D35171}{\textcolor{white}{\textbf{20}}} & \cellcolor[HTML]{AC2794}{\textcolor{white}{\textbf{31}}} & \cellcolor[HTML]{F9983E}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{49}}} & \cellcolor[HTML]{E26561}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{EB7557}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{F07F4F}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{28}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{F3874A}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{EF7C51}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{E66C5C}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{F79242}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{F07F4F}{\textcolor{white}{\textbf{2}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{rickshaw}}\\
-\hspace{1em}Never & 51 & 9 & 6 & 0 & 1 & 0\\
-\hspace{1em}Rarely & 64 & 70 & 18 & 4 & 3 & 1\\
-\hspace{1em}Once a week & 36 & 38 & 45 & 9 & 3 & 3\\
-\hspace{1em}2-3 per week & 34 & 36 & 18 & 29 & 6 & 4\\
-\hspace{1em}4-5 per week & 21 & 30 & 7 & 16 & 15 & 8\\
-\hspace{1em}Almost daily & 45 & 65 & 30 & 18 & 16 & 41\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{51}}} & \cellcolor[HTML]{EB7655}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{F3854B}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FBA139}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{340498}{\textcolor{white}{\textbf{64}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{70}}} & \cellcolor[HTML]{E26561}{\textcolor{white}{\textbf{18}}} & \cellcolor[HTML]{F99A3E}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FA9E3B}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{6001A6}{\textcolor{white}{\textbf{36}}} & \cellcolor[HTML]{4F02A2}{\textcolor{white}{\textbf{38}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{45}}} & \cellcolor[HTML]{F07F4F}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{2C0594}{\textcolor{white}{\textbf{34}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{36}}} & \cellcolor[HTML]{BE3885}{\textcolor{white}{\textbf{18}}} & \cellcolor[HTML]{6100A7}{\textcolor{white}{\textbf{29}}} & \cellcolor[HTML]{F89441}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{9511A1}{\textcolor{white}{\textbf{21}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{30}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{C8437B}{\textcolor{white}{\textbf{16}}} & \cellcolor[HTML]{D04D73}{\textcolor{white}{\textbf{15}}} & \cellcolor[HTML]{F99A3E}{\textcolor{white}{\textbf{8}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{99159F}{\textcolor{white}{\textbf{45}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{65}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{30}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{18}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{16}}} & \cellcolor[HTML]{AE2892}{\textcolor{white}{\textbf{41}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{CNG auto-rickshaw}}\\
-\hspace{1em}Never & 176 & 14 & 3 & 1 & 0 & 0\\
-\hspace{1em}Rarely & 213 & 95 & 14 & 4 & 1 & 3\\
-\hspace{1em}Once a week & 36 & 31 & 38 & 9 & 2 & 3\\
-\hspace{1em}2-3 per week & 29 & 18 & 16 & 14 & 8 & 3\\
-\hspace{1em}4-5 per week & 13 & 6 & 5 & 9 & 14 & 1\\
-\hspace{1em}Almost daily & 8 & 7 & 2 & 0 & 0 & 4\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{176}}} & \cellcolor[HTML]{F79044}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA537}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{213}}} & \cellcolor[HTML]{BE3885}{\textcolor{white}{\textbf{95}}} & \cellcolor[HTML]{F89441}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA438}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{280592}{\textcolor{white}{\textbf{36}}} & \cellcolor[HTML]{5901A5}{\textcolor{white}{\textbf{31}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{38}}} & \cellcolor[HTML]{E97158}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FA9E3B}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{29}}} & \cellcolor[HTML]{9D189D}{\textcolor{white}{\textbf{18}}} & \cellcolor[HTML]{B12A90}{\textcolor{white}{\textbf{16}}} & \cellcolor[HTML]{C23C81}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{E97257}{\textcolor{white}{\textbf{8}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{310597}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{CA457A}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{930FA3}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{8}}} & \cellcolor[HTML]{42039E}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{E16462}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{B12A90}{\textcolor{white}{\textbf{4}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{bus}}\\
-\hspace{1em}Never & 116 & 9 & 3 & 0 & 0 & 0\\
-\hspace{1em}Rarely & 155 & 46 & 9 & 0 & 0 & 1\\
-\hspace{1em}Once a week & 73 & 32 & 22 & 4 & 0 & 0\\
-\hspace{1em}2-3 per week & 43 & 19 & 11 & 7 & 5 & 0\\
-\hspace{1em}4-5 per week & 60 & 6 & 5 & 4 & 12 & 1\\
-\hspace{1em}Almost daily & 103 & 21 & 10 & 11 & 2 & 10\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{116}}} & \cellcolor[HTML]{F79044}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FA9E3B}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{155}}} & \cellcolor[HTML]{D9586A}{\textcolor{white}{\textbf{46}}} & \cellcolor[HTML]{F89540}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA438}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{73}}} & \cellcolor[HTML]{BE3885}{\textcolor{white}{\textbf{32}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{22}}} & \cellcolor[HTML]{F9973F}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{43}}} & \cellcolor[HTML]{BE3885}{\textcolor{white}{\textbf{19}}} & \cellcolor[HTML]{E06363}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{ED7A53}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{F3854B}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{60}}} & \cellcolor[HTML]{F68E45}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{F79442}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{F9983E}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{E97357}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{103}}} & \cellcolor[HTML]{E97357}{\textcolor{white}{\textbf{21}}} & \cellcolor[HTML]{F79044}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{F68C45}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{F79044}{\textcolor{white}{\textbf{10}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{motorcycle}}\\
-\hspace{1em}Never & 467 & 22 & 12 & 3 & 1 & 8\\
-\hspace{1em}Rarely & 44 & 38 & 13 & 10 & 0 & 4\\
-\hspace{1em}Once a week & 11 & 11 & 34 & 3 & 3 & 3\\
-\hspace{1em}2-3 per week & 13 & 6 & 6 & 15 & 4 & 2\\
-\hspace{1em}4-5 per week & 6 & 4 & 0 & 4 & 14 & 3\\
-\hspace{1em}Almost daily & 1 & 3 & 2 & 4 & 3 & 23\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{467}}} & \cellcolor[HTML]{F99A3E}{\textcolor{white}{\textbf{22}}} & \cellcolor[HTML]{FB9F3A}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{FCA537}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FBA238}{\textcolor{white}{\textbf{8}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{44}}} & \cellcolor[HTML]{46039F}{\textcolor{white}{\textbf{38}}} & \cellcolor[HTML]{DA5A6A}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{E56A5D}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{F68C45}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{DF6263}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{DF6263}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{34}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{4B03A1}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{15}}} & \cellcolor[HTML]{EF7C51}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{C13B82}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{E66C5C}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{F68C45}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{F9993E}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{F0804E}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{F68C45}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{23}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{walk}}\\
-\hspace{1em}Never & 17 & 7 & 3 & 4 & 0 & 2\\
-\hspace{1em}Rarely & 26 & 53 & 13 & 6 & 2 & 4\\
-\hspace{1em}Once a week & 12 & 19 & 54 & 9 & 4 & 2\\
-\hspace{1em}2-3 per week & 10 & 14 & 12 & 26 & 13 & 4\\
-\hspace{1em}4-5 per week & 7 & 3 & 7 & 11 & 16 & 8\\
-\hspace{1em}Almost daily & 46 & 73 & 40 & 49 & 27 & 197\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{17}}} & \cellcolor[HTML]{C43E7F}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{EB7655}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{E3685F}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}} & \cellcolor[HTML]{F3854B}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{B7318A}{\textcolor{white}{\textbf{26}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{53}}} & \cellcolor[HTML]{E66C5C}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{F79044}{\textcolor{white}{\textbf{6}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{E97257}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{D5536F}{\textcolor{white}{\textbf{19}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{54}}} & \cellcolor[HTML]{F1814D}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{DD5E66}{\textcolor{white}{\textbf{10}}} & \cellcolor[HTML]{BB3488}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{CD4A76}{\textcolor{white}{\textbf{12}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{26}}} & \cellcolor[HTML]{C5407E}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{D8576B}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{930FA3}{\textcolor{white}{\textbf{11}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{16}}} & \cellcolor[HTML]{CA457A}{\textcolor{white}{\textbf{8}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{F3874A}{\textcolor{white}{\textbf{46}}} & \cellcolor[HTML]{DE5F65}{\textcolor{white}{\textbf{73}}} & \cellcolor[HTML]{F79143}{\textcolor{white}{\textbf{40}}} & \cellcolor[HTML]{F1824C}{\textcolor{white}{\textbf{49}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{27}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{197}}}\\
 \addlinespace[0.3em]
 \multicolumn{7}{l}{\textbf{bicycle}}\\
-\hspace{1em}Never & 504 & 22 & 7 & 4 & 2 & 4\\
-\hspace{1em}Rarely & 59 & 43 & 20 & 4 & 2 & 4\\
-\hspace{1em}Once a week & 14 & 13 & 21 & 3 & 2 & 1\\
-\hspace{1em}2-3 per week & 3 & 4 & 5 & 7 & 2 & 0\\
-\hspace{1em}4-5 per week & 2 & 2 & 4 & 1 & 9 & 3\\
-\hspace{1em}Almost daily & 7 & 1 & 1 & 1 & 3 & 16\\
+\hspace{1em}Never & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{504}}} & \cellcolor[HTML]{FA9B3D}{\textcolor{white}{\textbf{22}}} & \cellcolor[HTML]{FCA238}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{FCA537}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FCA537}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}Rarely & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{59}}} & \cellcolor[HTML]{7501A8}{\textcolor{white}{\textbf{43}}} & \cellcolor[HTML]{D5546E}{\textcolor{white}{\textbf{20}}} & \cellcolor[HTML]{FA9C3C}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FA9C3C}{\textcolor{white}{\textbf{4}}}\\
+\hspace{1em}Once a week & \cellcolor[HTML]{8808A6}{\textcolor{white}{\textbf{14}}} & \cellcolor[HTML]{9714A1}{\textcolor{white}{\textbf{13}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{21}}} & \cellcolor[HTML]{F58B47}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{F9983E}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}}\\
+\hspace{1em}2-3 times per week & \cellcolor[HTML]{C13B82}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{9E199D}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{7601A8}{\textcolor{white}{\textbf{5}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{DB5C68}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{0}}}\\
+\hspace{1em}4-5 times per week & \cellcolor[HTML]{F1844C}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{F1844C}{\textcolor{white}{\textbf{2}}} & \cellcolor[HTML]{CB4679}{\textcolor{white}{\textbf{4}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{9}}} & \cellcolor[HTML]{E16462}{\textcolor{white}{\textbf{3}}}\\
+\hspace{1em}Almost daily & \cellcolor[HTML]{C6417D}{\textcolor{white}{\textbf{7}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{FCA636}{\textcolor{white}{\textbf{1}}} & \cellcolor[HTML]{F1814D}{\textcolor{white}{\textbf{3}}} & \cellcolor[HTML]{0D0887}{\textcolor{white}{\textbf{16}}}\\
 \bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}Darker cell colors indicate higher cell values in a row.}\\
 \end{tabular}}
@@
 The frequency of trip-making by other modes changed more noticeably: the
-frequency of travel by ridesharing services, rickshaw, CNG
+frequency of travel by ride-hailing services, rickshaw, CNG
 auto-rickshaw, and bus collapsed, with vastly more respondents reporting
@@
 this mode remained is low. In contrast, we see that the probabilities of
-never ridesharing during the pandemic are high irrespective of the
+never ride-hailing during the pandemic are high irrespective of the
 initial level of use of this mode of transportation, and the probability
@@
 \caption{\label{fig:circular-plot-1}Transition probabilities in trip
-frequency from before to during COVID-19: car and rideshare}
+frequency from before to during COVID-19: car and ride-hailing}
 \end{figure}
@@
 
-After rideshare, the use of bus had the largest probabilities of
+After ride-hailing, the use of bus had the largest probabilities of
 discontinuation (Figure \ref{fig:circular-plot-3}). We see that even
@@
 mobility across-the-board, but modes that require interaction with
-strangers (rideshare, bus, rickshaw, and CNG auto-rickshaw) were more
+strangers (ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more
 affected. Travel by car, motorcycle, and bicycle were somewhat less

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
shaila

--- 2020-10-06 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
@@
  The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
-Recent helps to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
+Recent studies help to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.
 The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
-In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw.
+In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw, and some adoption of walking during the pandemic.
 
@@
 Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
-The survey was available both in Bengali and English and disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
+The country was under partial lock-down during the survey period where essential services remained open, offices were running on rotation basis, all educational institutes were closed and public transport services were running at 50% capacity.
+The survey was disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
 In addition to socio-economic and demographic information of the respondents, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd

--- 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
@@
  The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
-Recent helps to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
+Recent studies help to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Response-to-Reviewers-v2.Rmd
v2 Continued drafting response to reviewers
Addressed reviewers comments

@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{Rows and columns are now clearly labelled as "Before" and "During", respectively. We have also color-coded the cells by rows, to more easily see the patterns from before to during.}
 
 - Otherwise, overall the paper is well written and the results are clearly discussed.
+
+\textcolor{blue}{Thank you.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{Thanks for asking this question. We clarified this in the third and fourth paragraph of section 'Methods and Data'.
+
+Here is what we included: Looking into the dataset, we can see that more than 80% of the sample belongs to the 16-30 years age group whereas 37% of the country's population belongs to this group @bangladesh2015.
+66% of the respondents are male and 32% are females compared to country's 49:51 male-female ratio @bangladesh2015.
+This is expected in terms of online surveys as young adults are more familiar with the online platforms, specially in the developing country context.
+Online surveys usually don't offer representative samples compared to face-to-face survey, rather a convenient sample @fricker2002advantages.
+However, considering the country-wide lock-down situation, the required physical distancing recommendation and relevant health regulations, we decided to conduct an online survey which is still better than having no data on travel behavior during COVID - 19 in Bangladesh.
+Also, in case of Bangladesh, collecting a representative sample even through face-to-face survey is very difficult because of overall reluctance in participation from public, along with women's strong unwillingness to share their information to unknown parties [see @jamal2020perceptions for more details].
+In terms of household's monthly income, 20% belongs to households with less than BDT 20,000, 25% belongs to BDT 20,000 - 39,999 and 22% belongs to BDT 40,000 - 59,999 income.
+[1 USD = 85 BDT]
+
+In terms of geography, we asked respondents to mention the name of the district they reside.
+48% of the respondents live in Dhaka.
+Although, there are some differences in availability of transportation services between cities, especially between Dhaka and cities outside Dhaka, the mode use frequency indicates that only 15% never used buses and 32% never used ride-hailing services before COVID - 19, which means these modes were available to most of the respondents and a vast majority has used them to some extent before the lock-down (see Figure \ref{fig:column-plot-cases}).
+As the total population of the country is 160 million, compared to the overall population, sample size is still small which is a limitation of the study.}
 
@@
 
+\textcolor{blue}{Thank you for this comment. We clarified this in the third and fourth paragraph of section 'Methods and Data'. 
+Here is what we included: Looking into the dataset, we can see that more than 80% of the sample belongs to the 16-30 years age group whereas 37% of the country's population belongs to this group @bangladesh2015.
+66% of the respondents are male and 32% are females compared to country's 49:51 male-female ratio @bangladesh2015.
+This is expected in terms of online surveys as young adults are more familiar with the online platforms, specially in the developing country context.
+Online surveys usually don't offer representative samples compared to face-to-face survey, rather a convenient sample @fricker2002advantages.
+However, considering the country-wide lock-down situation, the required physical distancing recommendation and relevant health regulations, we decided to conduct an online survey which is still better than having no data on travel behavior during COVID - 19 in Bangladesh.
+Also, in case of Bangladesh, collecting a representative sample even through face-to-face survey is very difficult because of overall reluctance in participation from public, along with women's strong unwillingness to share their information to unknown parties [see @jamal2020perceptions for more details].
+In terms of household's monthly income, 20% belongs to households with less than BDT 20,000, 25% belongs to BDT 20,000 - 39,999 and 22% belongs to BDT 40,000 - 59,999 income.
+[1 USD = 85 BDT]
+
+In terms of geography, we asked respondents to mention the name of the district they reside.
+48% of the respondents live in Dhaka.
+Although, there are some differences in availability of transportation services between cities, especially between Dhaka and cities outside Dhaka, the mode use frequency indicates that only 15% never used buses and 32% never used ride-hailing services before COVID - 19, which means these modes were available to most of the respondents and a vast majority has used them to some extent before the lock-down (see Figure \ref{fig:column-plot-cases}).
+As the total population of the country is 160 million, compared to the overall population, sample size is still small which is a limitation of the study.}
+
 Other comments:
@@
 
+\textcolor{blue}{Thanks for this comment.
+The country was under partial lock-down during the survey period where essential services remained open, offices were running on rotation basis, all educational institutes were closed and public transport services were running at 50% capacity.
+
+We included this information on the first paragraph of section 'Method and Data'}
+
 - Table 1: columns and rows need to indicate which one is pre and during COVID, it has to be added to the table.
+
+\textcolor{blue}{Rows and columns are now clearly labelled as "Before" and "During", respectively. Also, we color-coded the cells by rows, after a comment by Reviewer 1.}
 

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
Addressed reviewers comments

@@
 
-Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
-The country was under partial lock-down during the survey period where essential services remained open, offices were running on rotation basis, all educational institutes were closed and public transport services were running at 50% capacity.
+Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period", conducted during July - August 2020 in Bangladesh.
+The country was under partial lock-down during the survey period where essential services remained open, offices were running on a rotation basis, all educational institutes were closed and public transport services were running at 50% capacity.
 The survey was disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
-In addition to socio-economic and demographic information of the respondents, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
+In addition to the respondents' socio-demographic information, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
 
@@
 There are $n=800$ responses in the data set.
+
+Looking into the dataset, we can see that more than 80% of the sample belongs to the 16-30 years age group whereas 37% of the country's population belongs to this group @bangladesh2015.
+66% of the respondents are male and 32% are females compared to country's 49:51 male-female ratio @bangladesh2015.
+This is expected in terms of online surveys as young adults are more familiar with online platforms, especially in the developing country context.
+Online surveys usually don't offer representative samples compared to face-to-face survey, rather a convenient sample @fricker2002advantages.
+However, considering the country-wide lock-down situation, the required physical distancing recommendation and relevant health regulations, we decided to conduct an online survey, which is still better than having no data on travel behavior during COVID - 19 in Bangladesh.
+Also, in case of Bangladesh, collecting a representative sample even through face-to-face survey is very difficult because of overall reluctance in participation from public, along with women's strong unwillingness to share their information to unknown parties [see @jamal2020perceptions for more details].
+In terms of household's monthly income, 20% belongs to households with less than BDT 20,000, 25% belongs to BDT 20,000 - 39,999 and 22% belongs to BDT 40,000 - 59,999 income.
+[1 USD = 85 BDT]
+
+In terms of geography, we asked respondents to mention the name of the district they reside.
+48% of the respondents live in Dhaka.
+Although, there are some differences in availability of transportation services between cities, especially between Dhaka and cities outside Dhaka, the mode use frequency indicates that only 15% never used buses and 32% never used ride-hailing services before COVID - 19, which means these modes were available to most of the respondents and a vast majority has used them to some extent before the lock-down (see Figure \ref{fig:column-plot-cases}).
+As the total population of the country is 160 million, compared to the overall population, sample size is still small which is another limitation of the study.
 

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Response-to-Reviewers-v2.Rmd
v2 First version of revisions and response to reviewers for proof-reading

--- 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Response-to-Reviewers-v2.Rmd
@@
 
-\textcolor{blue}{Done. Hopefully the aspect of the figure will make it easier to read the results.}
+\textcolor{blue}{Done.
+Hopefully the aspect of the figure will make it easier to read the results.
+Please also note that separate figures are generated for publication purposes, and these have better resolution.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{Rows and columns are now clearly labelled as "Before" and "During", respectively.
+We have also color-coded the cells by rows, to more easily see the patterns of change from before to during by mode and starting frequency.}
 
 - Otherwise, overall the paper is well written and the results are clearly discussed.
+
+\textcolor{blue}{Thank you.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{Thank you for this question, which echoes a question by Reviewer 2.
+We clarified this in the section 'Methods and Data'.
+The discussion in the revised paper provides additional information regarding the sample and its characteristics.
+In particular, the data set is a convenience sample which, while cannot be said to be representative of the population, is more homogeneous for younger, urban respondents.
+Convenience samples are used to address quickly emerging questions (see Fricker and Schonlau, 2002 in Field Methods, and Jager et al., 2017, in Monographs of the Society for Research in Child Development).
+Convenience samples have been used in COVID-19 research (for instance by Astroza et al., 2020, in Transport Findings) to generate knowledge in the context of a rapidly evolving public health situation.
+Bangladesh, in particular, is a challenging country to obtain probabilistic samples, due to a general reticence of the public, and in particular females, to provide information to unknown parties, as discussed in more detail in Jamal et al.
+(2020) in Transportation Research Part D, among other authors (also see Mitra, 2016).
+Hopefully the revised version does a better job of explaining the context and limitations of the study.}
+
+\textcolor{blue}{Thank you again for your feedback, and we hope that the revisions will address your suggestions to your satisfaction.}
+
+\textcolor{blue}{Mitra, Suman Kumar, 2016.
+Land Use, Land Value, and Transportation: Essays on Accessibility, Carless Households, and Long-Distance Travel.
+PhD thesis, University of California, Irvine.}
 
@@
 
+\textcolor{blue}{Thank you for this comment.
+You are right that this is a convenience sample, something that was not explicitly discussed in the original submission, but that we have clarified in the section 'Methods and Data' of the revised paper.}
+
+\textcolor{blue}{Convenience samples have the advantage of being quick and cost-efficient and, despite their lower generalizability, when they are homogeneous they are useful to address rapidly emerging questions (see Fricker and Schonlau, 2002 in Field Methods, and Jager et al., 2017, in Monographs of the Society for Research in Child Development).
+As you suspected, this particular data set has an over-representation of the 16-30 years age group, whereas 37\% of the country's population belongs to this group.
+Also, 66\% of the respondents are male and 32\% are females compared to country's 49:51 male-female ratio.
+These characteristics of the sample are expected in terms of online surveys as young adults are more familiar with online platforms, especially in the context of a developing country context.
+Also, in the case of Bangladesh, collecting a representative sample remains difficult because of overall public reluctance to participate, with women's strong unwillingness to share their information to unknown parties in particular in face-to-face interviews (see Jamal et al., 2020 in Transportation Research Part D).
+Interestingly, compared to  Jamal et al.
+(2020), who had only 15\% of responses from females, the online survey used in the present research achieved greater penetration among women.}
+
+\textcolor{blue}{Given these characteristics, the sample cannot be said to be a probabilistic random sample, but is rather homogeneous for urban, younger respondents, and has greater representation from females than comparable samples obtained from face-to-face interviews in Bangladesh.
+We would argue that the benefits of reporting the information while clearly noting the characteristics of the convenience sample outweigh the limitations (also see Astroza et al., 2020, in Transport Findings).}
+
 Other comments:
@@
 
+\textcolor{blue}{Thanks for this comment.
+In the revised version we provide further information about this, and note that the country was under partial lock-down during the survey period, which meant that only essential services remained open, offices were running on rotation basis, all educational institutes were closed and public transport services were running at 50\% capacity.}
+
 - Table 1: columns and rows need to indicate which one is pre and during COVID, it has to be added to the table.
+
+\textcolor{blue}{Rows and columns are now clearly labelled as "Before" and "During", respectively. Also, we color-coded the cells by rows, after a comment by Reviewer 1.}
 
@@
 \textcolor{blue}{This has been changed throughout the paper.}
+
+\textcolor{blue}{Thank you again for your feedback, and we hope that the revisions will address your suggestions to your satisfaction. We believe that the paper has been improved as a result.}

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
v2 First version of revisions and response to reviewers for proof-reading

--- 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
@@
 ---
-title: Changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
+title: "Changes in trip-making frequency by mode during the COVID-19 emergency: evidence from an online survey in Bangladesh"
 author:
@@
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.
-The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
-In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw, and some adoption of walking during the pandemic.
+The analysis is based on an online sample conducted during the pandemic, and the results confirm an overall loss of mobility, especially among younger people, in the form of reduced trip-making frequency by all modes.
+In addition, the results suggest that changes are not uniform across modes, and in particular the loss of mobility was more pronounced for bus, rickshaw, and CNG auto-rickshaw.
+In contrast, there was some adoption of walking during the pandemic.
 
@@
 The objective of this paper is to investigate changes in the trip-making frequency by different modes of transportation during the COVID-19 emergency in Bangladesh.
-Using data from a recent survey that asked respondents to report mobility levels before and during the pandemic, we pose the following questions:
+Using data from a recent online survey that asked respondents to report mobility levels before and during the pandemic, we pose the following questions:
 
@@
 
-Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
-The country was under partial lock-down during the survey period where essential services remained open, offices were running on rotation basis, all educational institutes were closed and public transport services were running at 50% capacity.
+Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period", conducted during July - August 2020 in Bangladesh ($n=800$).
 The survey was disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
-In addition to socio-economic and demographic information of the respondents, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
+In addition to the respondents' socio-demographic information, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
 
+Like other surveys conducted during COVID-19 [e.g., @Astroza2020mobility], this is a convenience sample.
+Convenience samples have the advantage of being quick and cost-efficient and, despite their lower generalizability, when they are homogeneous they are useful to address rapidly emerging questions [Fricker2002advantages; @Jager2017convenience].
+This particular data set has an over-representation of the 16-30 years age group, whereas 37% of the country's population belongs to this group [@Bangladesh2015].
+Also, 66% of the respondents are male and 32% are females compared to country's 49:51 male-female ratio [@Bangladesh2015].
+These characteristics of the sample are expected in terms of online surveys as young adults are more familiar with online platforms, especially in the context of a developing country context.
+Also, in the case of Bangladesh, collecting a representative sample remains difficult because of overall public reluctance to participate, with women's strong unwillingness to share their information to unknown parties in particular in face-to-face interviews [see @Jamal2020perceptions].
+In fact, compared to  Jamal et al.
+[-@Jamal2020perceptions; Table 1], who had only 15% of responses from females, the online survey used in the present research achieved greater penetration among women.
+
+In terms of geography, respondents were asked to identify the name of the district they reside.
+The resulting sample has 48% of respondents who live in Dhaka.
+Although, there are some differences in availability of transportation services between cities, especially between Dhaka and cities outside Dhaka, the mode use frequency indicates that only 15% never used buses and 32% never used ride-hailing services before COVID-19, which means these modes were available to most of the respondents and a vast majority has used them to some extent before the lock-down.
+
+Given the above characteristics, the sample cannot be said to be a probabilistic random sample, but is rather homogeneous for urban, younger respondents, and has greater representation from females than comparable samples obtained from face-to-face interviews in Bangladesh.
+
+The context for the study was the country-wide emergency situation during COVID-19, with required physical distancing recommendations and relevant health regulations in place.
+Bangladesh was under partial lock-down during the survey period, with only essential services open, offices running on a rotation basis, all educational institutes closed, and public transport services running at 50% capacity.
 For the purpose of this paper, we use two questions that provide information about trip-making frequency by eight modes of transportation.
@@
 Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
-There are $n=800$ responses in the data set.
 
@@
 
-The findings reported here provide important changes in mobility in a developing country during COVID-19.
+The findings reported here suggest important changes in mobility in a developing country during COVID-19, which (given the characteristics of the convenience sample) may be more representative of younger, urban residents in Bangladesh.
 In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more affected.

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.tex
v2 First version of revisions and response to reviewers for proof-reading

@@
             pdfauthor={},
-            pdftitle={Changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh},
+            pdftitle={Changes in trip-making frequency by mode during the COVID-19 emergency: evidence from an online survey in Bangladesh},
             colorlinks=false,
@@
 
-  \title{Changes in trip-making frequency by mode during the COVID-19 emergency
-in Bangladesh}
+  \title{Changes in trip-making frequency by mode during the COVID-19 emergency:
+evidence from an online survey in Bangladesh}
     \author[McMaster University]{Shaila Jamal}
@@
   The COVID-19 pandemic has had a profound impact on mobility in every
-  country and region around the world. Recent helps to illuminate some of
-  the dimensions of change - however, the evidence is still scant in
-  developing countries. The objective of this paper is to present an
+  country and region around the world. Recent studies help to illuminate
+  some of the dimensions of change - however, the evidence is still scant
+  in developing countries. The objective of this paper is to present an
   exploratory analysis of the changes in the frequency of trip-making by
-  mode during the COVID-19 emergency in Bangladesh. The results confirm an
-  overall loss of mobility in the form of reduced trip-making frequency by
-  all modes, but the changes are not uniform across modes. In particular,
-  there was a greater loss of mobility for bus, rickshaw, and CNG
-  auto-rickshaw, and some adoption of walking during the pandemic.
+  mode during the COVID-19 emergency in Bangladesh. The analysis is based
+  on an online sample conducted during the pandemic, and the results
+  confirm an overall loss of mobility, especially among younger people, in
+  the form of reduced trip-making frequency by all modes. In addition, the
+  results suggest that changes are not uniform across modes, and in
+  particular the loss of mobility was more pronounced for bus, rickshaw,
+  and CNG auto-rickshaw. In contrast, there was some adoption of walking
+  during the pandemic.
   \end{abstract}
@@
 frequency by different modes of transportation during the COVID-19
-emergency in Bangladesh. Using data from a recent survey that asked
-respondents to report mobility levels before and during the pandemic, we
-pose the following questions:
+emergency in Bangladesh. Using data from a recent online survey that
+asked respondents to report mobility levels before and during the
+pandemic, we pose the following questions:
 
@@
 potential of travel mode change behavior in the post-lock-down and
-post-pandemic (COVID-19) period'' conducted during July - August 2020 in
-Bangladesh. The survey was available both in Bengali and English and
-disseminated through various electronic means along with a widespread
-social media campaign and promotion in various social media sites and
-groups. In addition to socio-economic and demographic information of the
-respondents, the survey covered different aspects of travel such as
-travel behavior before and during COVID-19, knowledge related to
-COVID-19, and opinions and perceptions regarding travel behavior after
-lock-down and post-pandemic situation.
+post-pandemic (COVID-19) period'', conducted during July - August 2020
+in Bangladesh (\(n=800\)). The survey was disseminated through various
+electronic means along with a widespread social media campaign and
+promotion in various social media sites and groups. In addition to the
+respondents' socio-demographic information, the survey covered different
+aspects of travel such as travel behavior before and during COVID-19,
+knowledge related to COVID-19, and opinions and perceptions regarding
+travel behavior after lock-down and post-pandemic situation.
 
-For the purpose of this paper, we use two questions that provide
-information about trip-making frequency by eight modes of
-transportation. The modes are \textbf{car}, \textbf{ride-hailing} (e.g.,
-Uber, Pathao), \textbf{rickshaw}, \textbf{CNG auto-rickshaw} (a
-rickshaw-like vehicle powered by compressed natural gas), \textbf{bus},
-motorcycle/scooter (hereafter just \textbf{motorcycle}),
-\textbf{walking}, and \textbf{bicycle} (there was an additional
-catch-all category \textbf{other} which we ignore here). Participants in
-the survey used the following levels to report their frequency of
-traveling by each mode both before and during COVID-19: \emph{Never},
-\emph{Rarely}, \emph{Once a week}, \emph{2-3 trips per week}, \emph{4-5
-trips per week}, \emph{Almost daily}. There are \(n=800\) responses in
-the data set.
+Like other surveys conducted during COVID-19 (e.g., Astroza et al.,
+2020), this is a convenience sample. Convenience samples have the
+advantage of being quick and cost-efficient and, despite their lower
+generalizability, when they are homogeneous they are useful to address
+rapidly emerging questions (Fricker2002advantages; Jager et al., 2017).
+This particular data set has an over-representation of the 16-30 years
+age group, whereas 37\% of the country's population belongs to this
+group (Bangladesh Bureau of Statistics, 2015). Also, 66\% of the
+respondents are male and 32\% are females compared to country's 49:51
+male-female ratio (Bangladesh Bureau of Statistics, 2015). These
+characteristics of the sample are expected in terms of online surveys as
+young adults are more familiar with online platforms, especially in the
+context of a developing country context. Also, in the case of
+Bangladesh, collecting a representative sample remains difficult because
+of overall public reluctance to participate, with women's strong
+unwillingness to share their information to unknown parties in
+particular in face-to-face interviews (see Jamal et al., 2020). In fact,
+compared to Jamal et al.~{[}(2020); Table 1{]}, who had only 15\% of
+responses from females, the online survey used in the present research
+achieved greater penetration among women.
+
+In terms of geography, respondents were asked to identify the name of
+the district they reside. The resulting sample has 48\% of respondents
+who live in Dhaka. Although, there are some differences in availability
+of transportation services between cities, especially between Dhaka and
+cities outside Dhaka, the mode use frequency indicates that only 15\%
+never used buses and 32\% never used ride-hailing services before
+COVID-19, which means these modes were available to most of the
+respondents and a vast majority has used them to some extent before the
+lock-down.
+
+Given the above characteristics, the sample cannot be said to be a
+probabilistic random sample, but is rather homogeneous for urban,
+younger respondents, and has greater representation from females than
+comparable samples obtained from face-to-face interviews in Bangladesh.
+
+The context for the study was the country-wide emergency situation
+during COVID-19, with required physical distancing recommendations and
+relevant health regulations in place. Bangladesh was under partial
+lock-down during the survey period, with only essential services open,
+offices running on a rotation basis, all educational institutes closed,
+and public transport services running at 50\% capacity. For the purpose
+of this paper, we use two questions that provide information about
+trip-making frequency by eight modes of transportation. The modes are
+\textbf{car}, \textbf{ride-hailing} (e.g., Uber, Pathao),
+\textbf{rickshaw}, \textbf{CNG auto-rickshaw} (a rickshaw-like vehicle
+powered by compressed natural gas), \textbf{bus}, motorcycle/scooter
+(hereafter just \textbf{motorcycle}), \textbf{walking}, and
+\textbf{bicycle} (there was an additional catch-all category
+\textbf{other} which we ignore here). Participants in the survey used
+the following levels to report their frequency of traveling by each mode
+both before and during COVID-19: \emph{Never}, \emph{Rarely}, \emph{Once
+a week}, \emph{2-3 trips per week}, \emph{4-5 trips per week},
+\emph{Almost daily}.
 
@@
 
-The findings reported here provide important changes in mobility in a
-developing country during COVID-19. In general there was a loss of
-mobility across-the-board, but modes that require interaction with
-strangers (ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more
-affected. Travel by car, motorcycle, and bicycle were somewhat less
-affected, and there is evidence of some adoption of walking during the
-pandemic.
+The findings reported here suggest important changes in mobility in a
+developing country during COVID-19, which (given the characteristics of
+the convenience sample) may be more representative of younger, urban
+residents in Bangladesh. In general there was a loss of mobility
+across-the-board, but modes that require interaction with strangers
+(ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more affected.
+Travel by car, motorcycle, and bicycle were somewhat less affected, and
+there is evidence of some adoption of walking during the pandemic.
 
@@
 doi:\href{https://doi.org/10.32866/001c.13489}{10.32866/001c.13489}
+
+\leavevmode\hypertarget{ref-Bangladesh2015}{}%
+Bangladesh Bureau of Statistics, 2015. Population monograph of
+bangladesh. Age-sex composition of bangladesh population.
 
@@
 doi:\href{https://doi.org/10.1016/j.ssci.2020.104872}{10.1016/j.ssci.2020.104872}
+
+\leavevmode\hypertarget{ref-Jager2017convenience}{}%
+Jager, J., Putnick, D.L., Bornstein, M.H., 2017. II. MORE than just
+convenient: THE scientific merits of homogeneous convenience samples.
+Monographs of the Society for Research in Child Development 82, 13--30.
+doi:\href{https://doi.org/10.1111/mono.12296}{10.1111/mono.12296}
+
+\leavevmode\hypertarget{ref-Jamal2020perceptions}{}%
+Jamal, S., Mohiuddin, H., Paez, A., 2020. How do the perceptions of
+neighborhood conditions impact active transportation? A study in
+rajshahi, bangladesh. Transportation Research Part D: Transport and
+Environment 87, 102525.
 

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
v2 Corrected missing reference after proof-reading by Shaila, and rendered README.Rmd

@@
 Like other surveys conducted during COVID-19 [e.g., @Astroza2020mobility], this is a convenience sample.
-Convenience samples have the advantage of being quick and cost-efficient and, despite their lower generalizability, when they are homogeneous they are useful to address rapidly emerging questions [Fricker2002advantages; @Jager2017convenience].
+Convenience samples have the advantage of being quick and cost-efficient and, despite their lower generalizability, when they are homogeneous they are useful to address rapidly emerging questions [@Fricker2002advantages; @Jager2017convenience].
 This particular data set has an over-representation of the 16-30 years age group, whereas 37% of the country's population belongs to this group [@Bangladesh2015].

### 2020-10-23 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.tex
v2 Corrected missing reference after proof-reading by Shaila, and rendered README.Rmd

@@
 generalizability, when they are homogeneous they are useful to address
-rapidly emerging questions (Fricker2002advantages; Jager et al., 2017).
-This particular data set has an over-representation of the 16-30 years
-age group, whereas 37\% of the country's population belongs to this
-group (Bangladesh Bureau of Statistics, 2015). Also, 66\% of the
+rapidly emerging questions (Fricker and Schonlau, 2002; Jager et al.,
+2017). This particular data set has an over-representation of the 16-30
+years age group, whereas 37\% of the country's population belongs to
+this group (Bangladesh Bureau of Statistics, 2015). Also, 66\% of the
 respondents are male and 32\% are females compared to country's 49:51
@@
 
+\leavevmode\hypertarget{ref-Fricker2002advantages}{}%
+Fricker, R.D., Schonlau, M., 2002. Advantages and disadvantages of
+internet research surveys: Evidence from the literature. Field methods
+14, 347--367.
+
 \leavevmode\hypertarget{ref-Huynh2020culture}{}%

### 2020-10-23 README.Rmd
v2 Corrected missing reference after proof-reading by Shaila, and rendered README.Rmd

--- 2020-10-06 Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/Frequency-of-Travel-by-Mode-COVID-19-Bangladesh.Rmd
@@
 ---
-title: An exploratory analysis of changes in trip-making frequency by mode during the COVID-19 emergency in Bangladesh
-author:
-  - name: Shaila Jamal
-    email: jamals16@mcmaster.ca
-    affiliation: McMaster University
-  - name: Antonio Paez
-    email: paezha@mcmaster.ca
-    affiliation: McMaster University
-    footnote: 1
-address:
-  - code: McMaster University
-    address: School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
+output: github_document
+always_allow_html: true
+bibliography: Frequency-of-Travel-by-Mode-COVID-19-Bangladesh/References.bib
+---
+
+<!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+```
+
+# Changes in trip-making frequency by mode during the COVID-19 emergency: evidence from an online survey in Bangladesh
+
+<!-- badges: start -->
+<!-- badges: end -->
+
+Shaila Jamal (jamals16@mcmaster.ca)  
+Antonio Paez (paezha@mcmaster.ca)  
+School of Geography and Earth Science  
+McMaster University  
+1280 Main Street West, Hamilton, ON L8S 4K1  
+Canada
+
+Paper submitted for peer review to Transport Findings
+
+## Abstract
+
 The COVID-19 pandemic has had a profound impact on mobility in every country and region around the world.
-Recent helps to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
+Recent studies help to illuminate some of the dimensions of change - however, the evidence is still scant in developing countries.
 The objective of this paper is to present an exploratory analysis of the changes in the frequency of trip-making by mode during the COVID-19 emergency in Bangladesh.
-The results confirm an overall loss of mobility in the form of reduced trip-making frequency by all modes, but the changes are not uniform across modes.
-In particular, there was a greater loss of mobility for bus, rickshaw, and CNG auto-rickshaw.
+The analysis is based on an online sample conducted during the pandemic, and the results confirm an overall loss of mobility, especially among younger people, in the form of reduced trip-making frequency by all modes.
+In addition, the results suggest that changes are not uniform across modes, and in particular the loss of mobility was more pronounced for bus, rickshaw, and CNG auto-rickshaw.
+In contrast, there was some adoption of walking during the pandemic.
 
-journal: "Transport Findings"
-date: "`r Sys.Date()`"
-bibliography: References.bib
-#linenumbers: true
-#numbersections: true
-csl: elsevier-harvard.csl
-output: rticles::elsevier_article
-always_allow_html: true
+## Keywords
 
-header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
----
+- COVID-19
+- Mobility
+- Trip-making frequency
+- Modes of transportation
+- Transition probabilities
+- Circular plots
 
@@
 
-Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period" conducted during July - August 2020 in Bangladesh.
-The survey was available both in Bengali and English and disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
-In addition to socio-economic and demographic information of the respondents, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
+Data used for this paper come from an online survey, "Exploring the potential of travel mode change behavior in the post-lock-down and post-pandemic (COVID-19) period", conducted during July - August 2020 in Bangladesh ($n=800$).
+The survey was disseminated through various electronic means along with a widespread social media campaign and promotion in various social media sites and groups.
+In addition to the respondents' socio-demographic information, the survey covered different aspects of travel such as travel behavior before and during COVID-19, knowledge related to COVID-19, and opinions and perceptions regarding travel behavior after lock-down and post-pandemic situation.
 
+Like other surveys conducted during COVID-19 [e.g., @Astroza2020mobility], this is a convenience sample.
+Convenience samples have the advantage of being quick and cost-efficient and, despite their lower generalizability, when they are homogeneous they are useful to address rapidly emerging questions [@Fricker2002advantages; @Jager2017convenience].
+This particular data set has an over-representation of the 16-30 years age group, whereas 37% of the country's population belongs to this group [@Bangladesh2015].
+Also, 66% of the respondents are male and 32% are females compared to country's 49:51 male-female ratio [@Bangladesh2015].
+These characteristics of the sample are expected in terms of online surveys as young adults are more familiar with online platforms, especially in the context of a developing country context.
+Also, in the case of Bangladesh, collecting a representative sample remains difficult because of overall public reluctance to participate, with women's strong unwillingness to share their information to unknown parties in particular in face-to-face interviews [see @Jamal2020perceptions].
+In fact, compared to  Jamal et al.
+[-@Jamal2020perceptions; Table 1], who had only 15% of responses from females, the online survey used in the present research achieved greater penetration among women.
+
+In terms of geography, respondents were asked to identify the name of the district they reside.
+The resulting sample has 48% of respondents who live in Dhaka.
+Although, there are some differences in availability of transportation services between cities, especially between Dhaka and cities outside Dhaka, the mode use frequency indicates that only 15% never used buses and 32% never used ride-hailing services before COVID-19, which means these modes were available to most of the respondents and a vast majority has used them to some extent before the lock-down.
+
+Given the above characteristics, the sample cannot be said to be a probabilistic random sample, but is rather homogeneous for urban, younger respondents, and has greater representation from females than comparable samples obtained from face-to-face interviews in Bangladesh.
+
+The context for the study was the country-wide emergency situation during COVID-19, with required physical distancing recommendations and relevant health regulations in place.
+Bangladesh was under partial lock-down during the survey period, with only essential services open, offices running on a rotation basis, all educational institutes closed, and public transport services running at 50% capacity.
 For the purpose of this paper, we use two questions that provide information about trip-making frequency by eight modes of transportation.
-The modes are **car**, **ridesharing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
+The modes are **car**, **ride-hailing** (e.g., Uber, Pathao), **rickshaw**, **CNG auto-rickshaw** (a rickshaw-like vehicle powered by compressed natural gas), **bus**, motorcycle/scooter (hereafter just **motorcycle**), **walking**, and **bicycle** (there was an additional catch-all category **other** which we ignore here).
 Participants in the survey used the following levels to report their frequency of traveling by each mode both before and during COVID-19: _Never_, _Rarely_, _Once a week_, _2-3 trips per week_, _4-5 trips per week_, _Almost daily_.
-There are $n=800$ responses in the data set.
 
@@
       bicycle_xtab) %>% 
-  kable("latex",
+  t() %>%
+  data.frame() %>% 
+  mutate(across(.cols = everything(), function(x) {
+    cell_spec(x, "html", bold = TRUE, color = "white", background = spec_color(x, end = 0.8, alpha = 0.3, option = "A", direction = -1))
+    })) %>%
+  t() %>%
+  data.frame() %>%
+  rename(Never = X1, Rarely = X2, `Once a week` = X3, `2-3 times per week` = X4, `4-5 times per week` = X5, `Almost daily` = X6) %>%
+  rownames_to_column(var = "Before") %>%
+  mutate(Before = rep(c("Never", "Rarely", "Once a week", "2-3 times per week", "4-5 times per week", "Almost daily"), 8)) %>%
+  kable("html",
         booktabs = TRUE,
+        escape = FALSE,
         caption = "\\label{tab:cross-tabulation} Cross-tabulation of counts of travel frequency by mode, before and during COVID-19") %>%
   kable_styling(latex_options = c("scale_down", "HOLD_position")) %>%
+  add_header_above(c(" " = 1, "During" = 6)) %>%
   pack_rows("car", 1, 6) %>%
-  pack_rows("rideshare", 7, 12) %>%
+  pack_rows("ride-hailing", 7, 12) %>%
   pack_rows("rickshaw", 13, 18) %>%
@@
 In contrast, only `r sum(bus_xtab[1, 2:6])` respondents, or `r paste(round(sum(bus_xtab[1, 2:6]) / sum(bus_xtab[,]) * 100, 2), "%", sep="")` of those who never used buses before started doing so during the pandemic.
-By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so - but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
+By way of comparison, `r paste(round(bicycle_xtab[6, 1] / sum(bicycle_xtab[6,]) * 100, 2), "%", sep="")` of respondents who cycled almost daily before the pandemic stopped doing so \- but `r paste(round(sum(bicycle_xtab[1, 2:6]) / sum(bicycle_xtab[,]) * 100, 2), "%", sep="")` who never cycled before started doing so during the pandemic.
 
@@
 circle.plot(bicycle_ptab2, "bicycle", nlines = -3)
+
+# Uncomment to save to jpg files for the publisher
+#dev.off()
 ```
 
-The findings reported here provide important changes in mobility in a developing country during COVID-19.
-In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (rideshare, bus, rickshaw, and CNG auto-rickshaw) were more affected.
+The findings reported here suggest important changes in mobility in a developing country during COVID-19, which (given the characteristics of the convenience sample) may be more representative of younger, urban residents in Bangladesh.
+In general there was a loss of mobility across-the-board, but modes that require interaction with strangers (ride-hailing, bus, rickshaw, and CNG auto-rickshaw) were more affected.
 Travel by car, motorcycle, and bicycle were somewhat less affected, and there is evidence of some adoption of walking during the pandemic.
